### 2023-10-30 paper/main.tex
Initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{graphicx}
\usepackage{subcaption}

\usepackage{amsfonts}
\expandafter\let\csname equation*\endcsname\relax

\expandafter\let\csname endequation*\endcsname\relax

\usepackage{amsmath}
\usepackage{listings}

\usepackage[most]{tcolorbox}

\definecolor{bg}{RGB}{255,249,227}
\definecolor{light_orange}{RGB}{255,193,134}

\newtcolorbox{optional}[1][]{%
enhanced jigsaw,
%oversize,
colback=bg,
boxrule=0pt,
overlay unbroken and first ={%
\draw[line width=0.2pt,double=bg,draw=bg!70!black,
    double distance=1pt,] (frame.north west) -- (frame.north east);
\draw[line width=0.2pt,double=bg,draw=bg!70!black,
    double distance=1pt,] (frame.south west) -- (frame.south east);},
breakable,
arc=0pt,outer arc=0pt,
#1}%

\newtcolorbox{tldr}[1][]{%
enhanced jigsaw,
%oversize,
colback=light_orange,
boxrule=0pt,
overlay unbroken and first ={%
\draw[line width=0.2pt,double=bg,draw=bg!70!black,
    double distance=1pt,] (frame.north west) -- (frame.north east);
\draw[line width=0.2pt,double=bg,draw=bg!70!black,
    double distance=1pt,] (frame.south west) -- (frame.south east);},
breakable,
arc=0pt,outer arc=0pt,
#1}%

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    % basicstyle=\ttfamily\footnotesize,
    basicstyle=\ttfamily,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
\lstMakeShortInline[columns=fixed]!
\lstset{style=mystyle}

%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\begin{document}

\title{A practical guide to Digital Micro-mirror Devices (DMDs) for wavefront shaping}

\author{Sébastien M. Popoff}
\address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
\ead{sebastien.popoff@espci.psl.eu}

\author{Rodrigo Gutiérrez-Cuevas}
\address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
\ead{rodrigo.gutierrez-cuevas@espci.psl.eu>}

\author{Yaron Bromberg}
\address{Racah Institute of Physics, The Hebrew University of Jerusalem, Israel}

\author{Maxime W. Matthès}
\address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}

% \address{IOP Publishing, Temple Circus, Temple Way, Bristol BS1 6HG, UK}
\vspace{10pt}
\begin{indented}
\item[]September 2023
\end{indented}

\begin{abstract}
Abstract
\end{abstract}

%
% Uncomment for keywords
%\vspace{2pc}
%\noindent{\it Keywords}: XXXXXX, YYYYYYYY, ZZZZZZZZZ
%
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
%
% Uncomment if a separate title page is required
%\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

\section{Introduction}

Since the advent of adaptive optics, various technologies have been employed 
to modulate the amplitude and/or phase of light. 
Early adaptive optics devices, utilized in fields like microscopy and astronomy, 
offer rapid modulation capable of compensating for the aberrations of optical systems 
in real-time. 
However, these devices are constrained by a limited number of actuators, 
restricting their utility in complex media where a large number of degrees of freedom is essential. 
Liquid Crystal Spatial Light Modulators (LC-SLMs), 
which allow for the control of light phase across typically more than a million pixels, 
have emerged as powerful tools for wavefront shaping in complex media 
since the seminal work of A. Most and I. Vellekoop in the mid-2000s~\cite{Vellekoop2007focusing}. 
Nonetheless, LC-SLMs are hampered by their slow response time, 
permitting only a modulation speed ranging from a few Hz to about 100 Hz. 
Digital Micromirror Devices (DMDs) have emerged as a technology bridging the gap 
between these two types of systems; 
they offer a large number of pixels (similar to LC-SLMs) and fast modulation speeds (typically up to several tens of kHz). 
However, DMDs are restricted to binary amplitude modulation and are not optimized for coherent light applications. 
Utilizing DMDs for coherent control of light in complex media is therefore non-trivial 
and necessitates specific adaptations for efficient use. 
This tutorial aims to provide a comprehensive guide 
for selecting and employing DMDs in wavefront shaping applications within complex media.\\

To comprehend both the capabilities and limitations
of DMD technology for coherent wavefront shaping,
it is crucial to understand the device's operating principles
and its original design intentions.
Investigated and developed by Texas Instruments since the late 1970s,
DMDs gained prominence in the 1990s for video projection applications
under the commercial name of Digital Light Processing (DLP)~\cite{Dudley2003emerging}.
The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
DMDs operate by toggling the state of small mirrors between two distinct angles, denoted as $\pm \theta_\text{DMD}$.
The device is originally engineered for amplitude modulation in video projection applications.
In this configuration, one mirror angle directs light into the projection lens,
while the alternate angle results in the light path being blocked (see Fig~\ref{fig:combined_pixel}).
Given that projectors utilize incoherent light and that the DMD plane is optically conjugated with the projection screen,
aberrations within the DMD plane are generally not problematic.
Similarly, phase fluctuations induced by temperature variations, 
as well as minor vibrations from the cooling hardware, are inconsequential in this context.
The DMD is designed to produce binary on/off modulation,
which is then leveraged to generate grayscale images via pulse-width modulation.
Color modulation is accomplished through the use of a color wheel in conjunction with a bright white light source.\\

Third-party companies have developped kits for research applications,
providing a DMD with a control board and a software interface. 
In particular, Vialux devices~\cite{}, provides a FPGA board that enable high speed modulation 
by allowing frames to be stored in the device memory. 
However, sandard Texas Instrument video projector evaluation modules can be converted into a wavefront shaping device~\cite{Cox2021converting}, 
albeit with a lower modulation speed.
Such systems can then be converted into phase or complex field modulators. 
It typically relies on encoding the optical phase into 
the spatial displacement of binaray spatial fringes displayed on the DMD 
and then filtering the high spatial frequencies in the Fourier plane~\cite{Lee1979binary}.
It allows multiple level complex modulation at the expense of a reduced spatial resolution.
The implementation and the performance of such a system are studied in a separate tutorial~\cite{RODRIGO} 
and will not be detailed here.
In the rest of the paper, we will assume that the DMD is used for complex modulation using such an approach.
While the present tutorial is not specific to any particular device, 
the software section will focus on the Vialux control board.
In a first part....\\

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
\centering
\includegraphics[width = \textwidth]{images/pixel_1.jpg}
\label{fig:pix_left}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
\centering
\includegraphics[width = \textwidth]{images/pixel_2.jpg}
\label{fig:pix_right}
\end{subfigure}
\caption{
\textbf{Principle of operation of a DMD in a digital projector.}
Left, incident light can be reflected towards the projection lens (state {\em on}), 
or onto a beam dump (state {\em off}).
Right, zoom on the pixels. 
Image adapted from \cite{JacksonDMD}.
}
\label{fig:combined_pixel}
\end{figure}

% \cite{JacksonDMD}

\section{Choosing the right DMD: Diffraction effects}

\subsection{1D model}

One important difference between liquid crystal modulators and DMDs 
concerns the geometry of the pixel surface. 
This leads to diffraction effects that can be detrimental 
to the modulation and efficiency of the system.
It strongly depends on the wavelength of the illumination, 
the pixel pitch, and the incident and outgoing angles. 
Thus, in addition to choosing an appropriate anti-reflection coating, 
it is important to check that the pixel pitch is well-suited for the configuration used.
Texas Instruments chips come with different possible pixel pitch $d$ 
ranging roughly from $5$ to $\sim25$\textmu m\cite{TI}.\\

Let's consider a 1D system,
as illustrated in Fig.~\ref{fig:grating_geom},  
% For the sake of simplicity and 
to qualitatively understand the problem.
% we place ourselves in the small angle approximation. 
This corresponds to incident and outgoing angles close to normal to the modulator surface 
and to a small tilt angle of the mirror. 
% We will give latter the final formulas in the 2D case.
We first assume that all the pixels are in the same state 
and are illuminated by a plane wave from the far field.
A pixelated modulator then acts exactly as a grating, 
with a period $d$ equal to the pixel pitch.
Modulators have a filling fraction limited by the hardware, 
typically around 90\%.
This corresponds to having an effective size of active pixels of $d' < d$.

In general, a grating leads to the observation of diffraction orders  
with varying intensities 
and angles $\theta_p$ given by the grating equation
$\text{sin}(\theta_p) = p\lambda/d = p \text{sin}(\theta_D)$,
where $\lambda$ is the wavelength of the light 
and $\theta_D$ the angle of the first order of diffraction.
The intensity of the diffraction orders depends on the size and the response of one pixel.
We see that we can decouple the effect of the periodicity of the grating, 
that acts on the angles of the orders of diffraction, 
and the response of one pixel, 
that acts on the envelope of the angular response.

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{images/grating_geom.pdf}
  \label{fig:flat_grating}

  \caption{
  \textbf{1D grating geometry.}
 ...
  }
  \label{fig:grating_geom}
  \end{figure}

For a normal incidence 
in the case of a LC-SLM, the response of one pixel is 
assumed constant over the pixel size. 
For a blazed grating, we also have a linear phase slope on each pixel due to the tilt $\theta_B$ of the pixels.
For an an arbitrary incidence angle $\alpha$, we also have a global phase slope 
that has the effect of trivially shifting the angular diffraction by an angle $\alpha$.\\

In essence, the incident angle $\alpha$ has the effect to shift the whole angular diffraction pattern 
with respect to the normal incidence case, 
while the blazed angle $\theta_B$, the tilt angle of the mirror in the DMD, 
has the effect to shift only the envelope by $2\theta_B$.
When the filling fraction is close to 100\%, 
i.e. $d\approx d'$,
the enveloppe for a flat grating is maximal 
at $\theta_0=-\alpha$ corresponding to the angle of the zero-th order,
and close to zero for the other orders.
We then observe a single diffraction order, 
which corresponds to the ideal case.
Adding a blazed angle $\theta_B$ leads to a shift of the envelope, 
the position of the maximum 
$\theta_0=2\theta_B-\alpha$ 
may not correspond anymore to a diffraction order.\\

\begin{optional}
  If we neglect the finite size of the device and illumination 
  and place ourselves in the small angle approximation, 
  we can write the field reflected off the device 
  for a plane wave illumination
  in the two systems as:

  \begin{equation}
  \begin{aligned}
    R_\text{flat}(x) &\propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
    R_\text{blazed}(x) &\propto 
        \underbrace{
          \left(
            \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
              \times 
              \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle} \\ \text{+ angle of incidence}}}
              %\text{blazed angle + angle of incidence}
          \right) 
        }_\text{pixel response}
        \otimes_x 
        \left[
        \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
        .
        \underbrace{
        e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
      }_\text{angle of incidence}
      \right]
      \, ,
  \end{aligned}
  \label{eq:grating_response}
  \end{equation}

  with $\Pi(x)$ the rectangular function,
  representing the finite size of the pixel, 
  defined as:

  \begin{equation}
    \Pi(x) = 
    \begin{cases}
    1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
    0, & \text{otherwise}.
    \end{cases}
    \end{equation}

  The intensity as a function of the angle in the far-field 
  is given, up to a homotetic transformation, 
  by the absolute value squared of 
  Fourier transform of Eq.~\ref{eq:grating_response}
  and reads:

  \begin{equation}
    \begin{aligned}
      I_\text{flat}(\theta) \propto 
        \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D) 
          \times 
        \text{sinc}^2\left( (\text{sin}(\theta)+\text{sin}(\alpha)) \frac{\lambda}{d'}\right)\\
        I_\text{blazed}(\theta) \propto 
        \underbrace{
          \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D) 
        }_\text{orders of diffraction}
        \times 
        \underbrace{
          \text{sinc}^2\left( (\text{sin}(\theta)+\text{sin}(2\theta_B-\alpha)) \frac{\lambda}{d'}\right)
         }_\text{envelope} \, .
    \end{aligned}
  \end{equation}

\end{optional}
A more accurate computation of the far field can be found  in ~\cite{Wang2023diffraction}.
We observe that the envelope (right hand term) is maximal for $\text{sin}(\theta_{max}) = \text{sin}(\alpha-2\theta_B)$ 
while the effect of the periodicity (left hand term) is maximal for $\text{sin}(\theta_p)+ \text{sin}(\alpha) = p \lambda/d$. 
with $p$ an integer value.
We represent in Fig.~\ref{fig:gratings} the angular response of a flat grating and a blazed grating 
for a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
For a flat grating, only the zero-th order has a significant and comparable intensity. 
For a blazed grating, in this example close to the worst case scenario, 
two diffraction orders have a significant intensity, 
and other orders also have a non-negligible contributions. 
% This implies that by selecting one order, we can loose more that 50\% of the light. 
% In a 2D system, we can then loose more that 75\% of the light.\\ 
The ideal case, i.e. when the maximum of the envelope corresponds to a diffraction order,
leading a single diffraction order containing most of the energy,
is obtained when the blazed grating equation is satistied~\cite{Casini2014on}:

\begin{equation}
  \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha) 
    = 2 \,\text{sin}(\theta_B)  \,\text{sin}(\theta_B-\alpha) 
    = p\frac{\lambda}{d} \, .
  \label{eq:blazed_eq}
\end{equation}

% Let's intereprete this qualitatively.
% For flat grating
% the incidence angle $\alpha$

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = \textwidth]{images/gratings_flat.pdf}
  \label{fig:flat_grating}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = \textwidth]{images/gratings_blazed.pdf}
  \label{fig:blazed_right}
  \end{subfigure}
  \caption{
  \textbf{Flat grating vs blazed grating.}
 ...
  }
  \label{fig:gratings}
  \end{figure}

% \begin{equation}
%     a = 
%   \left\{
%     \begin{aligned}
%         a\\b
% %         x + y   &= 7 \\
% %         5x - 2y &= -7
%     \end{aligned}
%   \right\}
% \end{equation}

% \begin{equation}
%   R_\text{flat}(x) \propto \Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d) 
% \end{equation}

% small angle approximation
% \begin{equation}
%   R_\text{blazed}(x) \propto \left[\Pi\left(x/d'\right) \times e^{j\frac{2\pi}{\lambda}\theta_B}\right]\otimes_x \sum_k \delta(x-k d) 
% \end{equation}

\subsection{2D case}

To analyze more precisely the effect of diffraction in a DMD, 
we need to consider the 2D surface of the modulator.
Let's define a cartesian coordinate system in the plane of the DMD
with axis $x$ and $y$ aligned with the pixel sides 
(see Fig.~\ref{fig:2d_geom}a.). 
Pixels are repeated along those directions. 
A technical difficulty is that the axis of rotation of the pixels 
is the diagonal of the pixels, 
hence rotated by 45 degrees with respect to the $x$ and $y$. 
For the alignment and the maniputlation of the optical setup
it is more convenient to work in incident and outgoing beams 
with optical axis contained in an horizontal plane. 
A simple and common solution consists in roatating the chip by 45 degrees
with respect to the horizontal plane 
such that the axis of rotation of the pixels is vertical.  
This configuration is illustrated in Fig.~\ref{fig:2d_geom}.

\begin{figure}
  \centering
  \includegraphics[width = 0.8\textwidth]{images/dmd_45.pdf}
  \caption{
  \textbf{2D grating geometry.}
 ...
  }
  \label{fig:2d_geom}
\end{figure}

The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions.
The incident angle $\alpha'$. 
Because of the 45 degrees angle between 
axis of rotation of the mirrors and axis, 
both 1D gratings have the same properties. 
A more complete computation of the 2D system can be found in~\cite{Scholes2019structured}.
To apply Eq.~\ref{eq:blazed_eq}, 
we first need to project the 
the different angles of the problem in the incident planes 
of the two 1D gratings.
It gives 
$\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$, 
$\beta = \text{arctan}\left(\text{tan}(\beta')/\sqrt{2}\right)$, 
and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$, 
$theta_\text{DMD}$ being the actual rotation angle of the mirrors with respect to the diagonal of the pixels.
We can quantify how close we are to the ideal case, 
i.e. when satifying the blazed equation~\ref{eq:blazed_eq}, 
by defining a {\em blazed number} $\mu$ as~\cite{WFSnet_diffraction}:

\begin{equation}
  \mu =  
  \left|
    \lfloor 2 \frac{d}{\lambda} 
      \left[
        \text{sin}(\alpha)  +\text{sin}(\beta)
      \right] 
    \rfloor
    \mod{2} -1 
  \right| \, ,
\end{equation}

with $\lfloor . \rfloor$ representing the integer part, 
and $\mod{2}$ the modulo 2 operation.
$\mu$ is maximal when the blazing equation is satisfield, 
i.e. when one order of diffraction contains most of the energy, 
and minimal when we are in the worst case scenario, 
i.e. when four orders of diffraction have a significant and equal intensity.\\

To illustrate the effect, we simulate a DMD in Python 
(see tutorial and code in~\cite{WFSnet_diffraction})
with a two pixel pitches of 
$d=7.6$\textmu m and $d=10.8$\textmu m,
for a coherent excitation at $\lambda=633$nm. 
We show in Fig.~\ref{fig:mu76} the blazed number $\mu$
as a function of the incident angle $\alpha'$ 
as well as the far field diffraction pattern for different two incident angles.
While one can change the efficiency of diffraction by tuning 
the incindent angle, its effect is somewhat limited 
withing a reasonable angular range compatible with experimental constraints 
(i.e. for angles far from $\pm 90^\circ$).
Far field pattern are centered around the maximum of the envelope 
$\theta_\text{max} = 2\theta_B - \alpha$ 
(marked by a yellow cross).
We see that for small values of $\alpha$, 
the pixel pitch of $d=10.8$\textmu m leads to a blazed number $\mu$ 
close $1$. 
It corresponds in the far field to having one bright order of diffraction  
close to the maximum of the envelope.

\begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = \textwidth]{images/mu_76.pdf}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = \textwidth]{images/mu_108.pdf}
  \end{subfigure}
  \caption{
  \textbf{Blazed number and far field diffraction pattern for...}
 ...
  }
  \label{fig:mu76}
\end{figure}

\subsection{Modulation cross-talk}

In practice, we place a pinhole or iris to select 
one order of diffraction,
corresponding to the {\em on} state.
Having a small value for the blaze number $\mu$
does not simply limit the amount of light modulated 
due to the reduced diffraction efficiency, 
it also impact the modulation qualoty by introducing cross-talk 
between the two states of the DMD pixels.
For now, we considered that all the pixels are in the same state. 
In a practical use of the DMD,
we need to modulate the state of each pixel individually.
When $\mu$ is close to zero, higher order of diffraction 
still have a non negligible intensity, 
as shown in the 1D case in Fig~\ref{fig:gratings}. 
One consequence is that pixels in the {\em off} state 
may have orders of diffraction that are not blocked by the pinhole, 
and will contribute as a perturbation to the modulated wavefront. 
While this contrinution may seem small, 
it does impact the quality of the modulation 
as the modulation scheme usually requires about half the pixels 
to be in the {\em off}  state for phase modulation~\cite{}, 
and even more for complex modulation schemes~\cite{}.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/xtalk.pdf}
  \caption{
  \textbf{Cross-talk between {\em on} and {\em off} states.}
 ...
  }
  \label{fig:xtalk}
\end{figure}
% For an incident angle $\alpha'$ 
% and a reflected angle $\beta'$ 
% in the horizontal plane,
% we can use the grating equation 
% presented in Eq.~\ref{eq:blazed_eq}

\begin{tldr}
  \textbf{TL;DR:}\\
  DMDs act as a blazed grating. 
  For a given operation wavelength,
  we need to find the right pixel pitch 
  to have a good modulation quality 
  and diffraction efficiency.
  \end{tldr}

% Let's consider first a 1D toy model to have a qualitative understanding of the problem.

% Liquid crystal pixels are flat and aligned with the array surface. 
% The periodicity of the pixels makes the device behave as a flat grating. 
% Thus, the reflection angle is the opposite of the incident one compared to the normal of the array plane. 
% That implies that the path coming from each element of the device has the same optical path length 
% (see Fig.~\ref{fig:reflection} left).

% In a DMD, the modulation is achieved by setting each pixel to one of two states, 
% corrsponding to angles $\pm \theta$ with respect to the the normal of the array plane. 
% It thus acts as a blazed grating.
% Two consecutive pixels, or elements, now give rise to a phase difference 
% between two parallel rays reflected of neighboring pixels (see Fig.~\ref{fig:reflection} right). 
% In the Fourier plane, where we commonly select the zero-th order, all the diffraction orders are shifted compared to the optical axis within the same envelope. In the worst scenario, we can end up with four diffraction orders with approximately the same intensity, each significantly shifted compared to the optical axis. An example is shown in Figure 6. This is almost exactly what I had the first time I set up my DMD while I expected something like in Figure 5.

% Another way of seeing this effect is to consider that the diffraction orders (satisfying the grating equation) do not match anymore the angle of the reflected beam.

% We first use a 1D model to evaluate the effect of the tilt angle on the diffraction pattern. We then use simple numerical calculations to estimate the profile of the diffraction pattern. The full Python code used in this post is available as an IPython Notebook on my GitHub account: blazing_angle_DMD.ipynb.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% ABERRATIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Characterizing aberration effects}

\subsection{Presentation of the problem}

While only able to provide a hardware binary amplitude modulation,
DMDs are a powerful tool for wavefront shaping and sensing. 
For these applications, 
aberrations due to the non-flatness of the DMD surface 
are detrimental and need to be characterized and corrected for. 
For LC-SLMs, inhomoegeneities of the surface is usually characterized by the 
manufacturer and the spatial phase profile of the introduced aberrations 
in the plane of the modulator is provided for the working wavelength.
As previously mentioned,
when used for intensity modulation in a plane conjugated to the DMD plane, 
as used in digital projectors, 
the system is not sensitive to aberrations brought by the DMD surface.
For these reasons, this effect is usually not documented, 
and usually not mentioned, in the information provided by the manufacturer 
despite the fact that the deviation from a flat surface
typically reaches few wavelengths~\cite{Brown2021multicolor}.
\\

\subsection{Finding the correction pattern}

In the litterature, different methods have been developped to characterize 
the phase pattern representing the aberrations of the DMD in the 
plane of the modulator. 
Typically, it consists in using a model of the aberations 
and adjusting the parameters to fit the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
While providing accurate results, 
the implementation of the approach has to be adapted to the specific 
setup condition and may require
additional optical components, 
precise alignments, 
and custom software. 
An other approach consists directly measuring the distorded wavefront, 
either using wavefront sensor, such as a Shack-Hartmann, 
or using interferometry. 
Again, it requires specific hardware and 
a carefully calibrated and stable setup.
In this section, we present a simple method to characterize the aberrations using a lens and a camera. 
This approach can be used for any system prividing a phase modulution, 
such as LC-SLMs or deformable mirrors.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/DMD_abberations_setup.pdf}
  \caption{
  \textbf{Setup for aberration correction.}
 ...
  }
  \label{fig:dmd_aberr_setup}
\end{figure}

He assume that the DMD is configured to provide a phase modulation~\cite{Lee1979binary, RODRIGO}. 
It implies that the modulator can be devided into regions, we call {\em macropixels},
on which the phase can be independently controlled.
We place a lens and a camera in its Fourier plane and illuminate the modulator with a collimated beam 
covering the full desired area of the modualtor we intend to use. 
If there is no or negligible aberrations, the observed intensity pattern 
reseembles the PSF of the lens, i.e. an Airy disk such as presented in Fig.~\ref{fig:dmd_aberr_setup}.c.
However, in partice, we observe a significantly distorted pattern,
such as the one presented in Fig.~\ref{fig:dmd_aberr_setup}.d.\\

We assume the the aberrations introduced by the DMD is smooth and 
can be described by a phase pattern $\phi^\text{aber}$ 
that can be well approximated by a finite number of 
Zernike polynomials $Z_n(r,\theta)$~\cite{Zernike1934beugungstheorie}, such as:

\begin{equation}
  \phi^\text{aber}(r,\theta) \approx \sum_{n=0}^N a_n Z_n(r,\theta) \, .
\end{equation}

The goal is to find and display the phase pattern 
on $\phi_i^\text{corr}$ on each pixel $i$ that best compensate for the aberrations, 
i.e. $\phi_i^\text{corr} = -\phi^\text{aber}(r_i,\theta_i)$.
We create this pattern in the basis of Zernike polynomials

\begin{equation}
  \phi_i^\text{corr} = \sum_{n=0}^N a'_n Z_n(r,\theta) \, ,
\end{equation}

the best correction if obtained for

\begin{equation}
  a'_n = -a_n \,\, \forall n \in [0..N]\, .
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/Zernike_1.pdf}
  \caption{
  \textbf{Correction phase map.}
 ...
  }
  \label{fig:phase_corr}
\end{figure}

We will perform a sequential optimization of the parameters $a_n$ 
to maximize a cost function that is optimial for an ideal correction of the abberations.
To define such a cost function, we first create a mask with a disk 
centered around the maximum of intensity of the initial image (resembling Fig.~\ref{fig:dmd_aberr_setup}.d)
and of radius equivalent to a speckle grain. 
This size does not need to be very very precisely set for the optimization procedure to work, 
and can be found by estimating the size of the ideal PSF, i.e. $r_0 \approx M \frac{\lambda}{2 NA}$, 
with $NA$ the numerical aperture of the optical system and $M$ its magnification.
For a given output image,  we compute the element-wise product between the intensity image measured and the mask and perform a summation.
We start with $a'_n = 0\,\, \forall n \in [0..N]$.
For each parameter, we try different values of $a'_n$,
generate the phases on each pixel so that 
$\phi_i^\text{corr} = \sum_{n=0}^N a'_n Z_n(r_i,\theta_i)$, 
measure the output intensity profile,
and estimate the value the cost function.
We keep for each parameter the value that leads to the highest value.
We repeat the whole process $3$ times for each parameter to 
mitigate the effect of noise or vibration.

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/Zernike_2.pdf}
  \caption{
  \textbf{Effect of Zernike polynomials on the optimizaton of the PSF.}
 ...
  }
  \label{fig:zernike}
\end{figure}

\subsection{Python code example}

We provide in the section Python pieces of code to perform the aberration correction.
We assume that the phase modulation can be controlled using a function 
!display_phase_mask(phase_pattern)! to display a desired phase pattern 
containing real values between $2\pi$.
We remind that phase modulation using a DMD requires an appropirate setup 
and procedure for the generation of the binary pattern that is not detailed here 
but detailed in ths tutorial~\cite{RODRIGO}.
We also assume a function !get_camera_image()! 
that returns the intensity pattern measured by the camera. 
We use the package !aotools!~\cite{Townson2019aotools} for generating Zernike polynomials.

% \lstinputlisting[language=Python, caption=Python example]{code/dmd_aberrations.py}

measuring projection of a signal wavefront with a displayed mask and a bucket detector~\cite{}

\cite{Lee1979binary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% 4. STABILITY %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Mechanical and thermal stability}

Unlike the original purpose of the DMD, i.e. ampltiude modulation for video projectors, 
typical scientific applications require a high stability of the generated wavefront. 
This is particularly true for applications in complex media, 
such as strongly scattering media or multimode fiber, 
where a small change in the phase front can lead to a large change in the output intensity profile. 
While LC-SLMs design has been improved and adapted to scientific applicaitons 
over the last decades, 
DMD are still a relatively new tools for wavefront shaping and sensing 
and are prone to instabilities that need to be adressed by the user.
In this section we present the effect of mechanical and thermal instabilities 
and how to limit their impact on the wavefront quality with simple solutions.\\

\subsection{Mechanical stability}

Most DMD kits are composed of two elements, 
the chip itself and the electronics board to control it. 
It can be the standard electronics board for video projectors,
as it is the case for TI evluations kits, 
or a FPGA for fast scientific use, 
as provided by Vialux~\cite{vialux}. 
These electronics come with a fan to cool the chip and the electronics board. 
As the connection between the chip and the electronics board 
used a rigid flat cable, 
the two elements are not mechanically connected.
Vibrations from the board are then transmitted to the chip,
leading to small angle variations of the mirror surface. 
While this effect is without consequences for video projection, 
it can be dramatic applications to complex media 
as their response is highly sensitive to the phase front.\\

Because of this high sensitivy, and since we want to use the DMD 
in the context of complex media, it is simpler to characterize this effect 
directly on the system response rather than 
building a different setup to characterize the wavefront itself.
We show an example of such setup in Fig.~\ref{fig:MMF_ref}. 
We expand an laser beam onto the DMD and sent the input light through a multimode fiber. 
While we typically want phase modulation, 
the setup will need additional elements to achieve this task~\cite{}. 
We here simplify the setup for the sake of clarity. 
The output of the fiber interfere with a reference arm in order to measure 
changes of the field.
We then measure the interference pattern using a camera. \\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/MMF_ref.pdf}
  \caption{
  \textbf{Measuring stability at the output of a multimode fiber.}
 ...
  }
  \label{fig:MMF_ref}
\end{figure}

We show in the supplementar maeterials~\cite{} an animation of the dynamic pattern. 
We also estimate the phase variation as a function of time. 
In a given region of the image, 
the relation between the phase and the transverse displacement of the fringes
is proportional to the phase, 
with a displacement equal to the fringes period corresponding to $2\pi$.
We show in Fig.~\ref{fig:phase_vibrations} (red curve) 
the fluctuation of the phase over time presenting fast oscillations
due to the mechanical vribration communicated by the board.\\

\begin{figure}
  \centering
  \includegraphics[width = 0.75\textwidth]{images/phase_vibrations.pdf}
  \caption{
  \textbf{Vibration indiced phase fluctuations.}
 ...
  }
  \label{fig:phase_vibrations}
\end{figure}

A simple but efficient solution consists in dumping the vibrations
at the level of the flat cable by clamping the cable with soft maeterial, 
as illustrated in Fig.~\ref{fig:dumping}. 
It can be done with off the shelf material.
We use here simple foam material used for packaging clamped to the cable 
using two metallic plates, screws, and nuts.
We observe a significant reduction of the phase fluctuations 
as shown in Fig.~\ref{fig:phase_vibrations} (blue curve).\\

\begin{figure}
  \centering
  \includegraphics[width = 0.95\textwidth]{images/dumping.pdf}
  \caption{
  \textbf{Dumping mechanical vibrations.}
 ...
  }
  \label{fig:dumping}
\end{figure}

\subsection{Thermal stability}

Electronics used to control the DMD chip heat up during operation.
This effect depends on the frame rate. 
In particular, it is important to notice that this effect is important when 
a sequence is running on the DMD, and is limited when the DMD is powered 
but without any sequence running.
Changes of temperature has the effect of deforming the surface of the chip 
and potentially changing the phase introduced by the glass protecting window.
This has the effect of introducing low order aberrations that degrade the quality of the wavefront.
While this effect is relatively less important than the ones due to 
static aberrations and mechanical instabilities detailed previously, 
it still have a significant impact on the response of a complex medium 
when using a DMD to modulate the input wavefront.\\

The first step consists in characterizing the effect of temperature. 
While a precise effect on the wavefront itself can be measured~\cite{Rudolf2021thermal}, 
it is once again usually more convenient and relevant to measure the effect directly on the response of the system studied.

Thermal stability~\cite{Rudolf2021thermal}

\begin{equation}
C(t) = 
  \frac{
    \left\langle 
     \bar{I}(\vec{r}, t) \bar{I}(\vec{r}, t=0) 
     \right\rangle_{\vec{r}}
  }{
    \sqrt{
     \left\langle 
      \bar{I}(\vec{r}, t)^2 
     \right\rangle_{\vec{r}}
     \left\langle 
      \bar{I}(\vec{r}, t=0)^2 
     \right\rangle_{\vec{r}}
    }
  }
\end{equation}

\begin{figure}[ht]
  \centering
  \includegraphics[width = 0.95\textwidth]{images/DMD_decorrelation_T.pdf}
  \caption{
  \textbf{Effect of temperature.}
    ...
  }
  \label{fig:Cizmar_1}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = 0.65\textwidth]{images/Cizmar_1.pdf}
  \caption{
  \textbf{Thermal stabilization of the DMD chip.}
    Image adapted from~\cite{Rudolf2021thermal}.
  }
  \label{fig:decorr_T}
\end{figure}

\begin{tldr}
\textbf{TL;DR:}\\
DMDs take about an hour to thermaly stabilize when a sequence is running. 
It can conuntered by using a thermoelectric cooler 
and a temperature sensor. 
It can also be simply mitigated by letting a sequence running for few hours
before starting the experiment. 
\end{tldr}

\section{Control Vialux DMDs with Python}

\cite{hofling2004alp}

\cite{popoff2016alp4lib}

\section{Acknowledgments}
Authors wishing to acknowledge assistance or encouragement from 
colleagues, special work by technical staff or financial support from 
organizations should do so in an unnumbered `Acknowledgments' section 
immediately following the last numbered section of the paper. In \verb"iopart.cls" the 
command \verb"\ack" sets the acknowledgments heading as an unnumbered
section.

\section{Bibliography}

\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}

### 2023-11-15 paper/main.tex
update text

@@
+\documentclass[12pt]{iopart}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-\documentclass[12pt]{iopart}
+
+% !TEX root =main.tex
+
 \usepackage{graphicx}
@@
 
-\usepackage[most]{tcolorbox}
+\usepackage[most, breakable]{tcolorbox}
+\usepackage{hyperref}
+
+\hypersetup{
+    breaklinks = true,
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,      
+    urlcolor=blue,
+    citecolor=red
+}
 
@@
     double distance=1pt,] (frame.south west) -- (frame.south east);},
-breakable,
 arc=0pt,outer arc=0pt,
@@
 \lstset{style=mystyle}
+
+\newcommand{\be}{\begin{equation}}
+\newcommand{\ee}{\end{equation}}
 
@@
 and necessitates specific adaptations for efficient use.
-This tutorial aims to provide a comprehensive guide 
-for selecting and employing DMDs in wavefront shaping applications within complex media.\\
+While other articles exist describing the various aspects of DMDs~\cite{Park2015properties, Cox2021converting, Wang2023diffraction},
+this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
+In particular, we provide characterization and validation procedures that requires minimal changes compared to typical wavefront shaping setups.\\
+
+REVIEW en bio imaging~\cite{Zhuang2020}
 
@@
 In this configuration, one mirror angle directs light into the projection lens,
-while the alternate angle results in the light path being blocked (see Fig~\ref{fig:combined_pixel}).
+while the alternate angle results in the light path being blocked (see Fig.~\ref{fig:combined_pixel}).
 Given that projectors utilize incoherent light and that the DMD plane is optically conjugated with the projection screen,
@@
 
-Third-party companies have developped kits for research applications,
-providing a DMD with a control board and a software interface. 
-In particular, Vialux devices~\cite{}, provides a FPGA board that enable high speed modulation 
-by allowing frames to be stored in the device memory. 
-However, sandard Texas Instrument video projector evaluation modules can be converted into a wavefront shaping device~\cite{Cox2021converting}, 
-albeit with a lower modulation speed.
-Such systems can then be converted into phase or complex field modulators. 
-It typically relies on encoding the optical phase into 
-the spatial displacement of binaray spatial fringes displayed on the DMD 
-and then filtering the high spatial frequencies in the Fourier plane~\cite{Lee1979binary}.
-It allows multiple level complex modulation at the expense of a reduced spatial resolution.
-The implementation and the performance of such a system are studied in a separate tutorial~\cite{RODRIGO} 
-and will not be detailed here.
-In the rest of the paper, we will assume that the DMD is used for complex modulation using such an approach.
-While the present tutorial is not specific to any particular device, 
-the software section will focus on the Vialux control board.
-In a first part....\\
+Third-party companies have developed kits tailored for research applications,
+which include a DMD, a control board, and a software interface.
+Specifically, Vialux devices~\cite{vialux} offer an FPGA board that enables high-speed modulation
+by allowing frames to be stored in the device's memory~\cite{hofling2004alp}.
+However, standard Texas Instruments video projector evaluation modules can also be repurposed into wavefront shaping devices~\cite{Cox2021converting},
+though at a compromised modulation speed.
+These systems can further be converted into phase or complex field modulators.
+This is typically achieved by encoding the optical phase into
+the spatial displacement of binary spatial fringes displayed on the DMD,
+followed by filtering the high spatial frequencies in the Fourier plane~\cite{lee1979binary}.
+Such a configuration permits multi-level complex modulation but sacrifices spatial resolution.
+The implementation and performance of such systems are explored in a separate tutorial~\cite{Gutierrez2024DMD}
+and will not be elaborated upon here.
+For the remainder of this paper, it will be assumed that the DMD is used for complex modulation via such a method.
+While the present tutorial is not constrained to any specific device,
+the software section will predominantly focus on the Vialux control board.
+In the first part....\\
 
@@
 
-% \cite{JacksonDMD}
-
 \section{Choosing the right DMD: Diffraction effects}
@@
 
-One important difference between liquid crystal modulators and DMDs 
-concerns the geometry of the pixel surface. 
-This leads to diffraction effects that can be detrimental 
-to the modulation and efficiency of the system.
-It strongly depends on the wavelength of the illumination, 
-the pixel pitch, and the incident and outgoing angles. 
-Thus, in addition to choosing an appropriate anti-reflection coating, 
-it is important to check that the pixel pitch is well-suited for the configuration used.
-Texas Instruments chips come with different possible pixel pitch $d$ 
-ranging roughly from $5$ to $\sim25$\textmu m\cite{TI}.\\
-
-Let's consider a 1D system,
-as illustrated in Fig.~\ref{fig:grating_geom},  
-% For the sake of simplicity and 
-to qualitatively understand the problem.
-% we place ourselves in the small angle approximation. 
-This corresponds to incident and outgoing angles close to normal to the modulator surface 
-and to a small tilt angle of the mirror. 
-% We will give latter the final formulas in the 2D case.
-We first assume that all the pixels are in the same state 
-and are illuminated by a plane wave from the far field.
-A pixelated modulator then acts exactly as a grating, 
-with a period $d$ equal to the pixel pitch.
-Modulators have a filling fraction limited by the hardware, 
+A significant distinction between liquid crystal modulators and DMDs
+lies in the geometry of the pixel surface.
+This difference gives rise to diffraction effects that can adversely affect
+both the modulation quality and system efficiency.
+The impact of these diffraction effects is highly contingent on several factors:
+the wavelength of the illumination, the pixel pitch, and both the incident and outgoing angles.
+Therefore, alongside selecting an appropriate anti-reflection coating,
+it is crucial to ensure that the pixel pitch is compatible with the specific configuration being used.
+Texas Instruments offers chips with a variety of pixel pitches $d$,
+ranging approximately from $5$ to $\sim25$ \textmu m~\cite{TI}.\\
+
+To achieve a qualitative understanding of this issue,
+we consider a 1D array of pixels as illustrated in Fig.~\ref{fig:grating_geom}.
+% For simplicity, we will focus on scenarios where both the incident and outgoing angles 
+% are nearly perpendicular to the modulator surface, and the mirror tilt angle is minimal.
+Initially, let's assume that all pixels are in the same state and are illuminated by a plane wave originating from the far field.
+Under these conditions, the pixelated modulator essentially functions as a grating,
+with a period $d$ that is equivalent to the pixel pitch.
+It is important to underscore that these modulators possess a hardware-limited fill factor,
 typically around 90\%.
-This corresponds to having an effective size of active pixels of $d' < d$.
-
-In general, a grating leads to the observation of diffraction orders  
-with varying intensities 
-and angles $\theta_p$ given by the grating equation
-$\text{sin}(\theta_p) = p\lambda/d = p \text{sin}(\theta_D)$,
-where $\lambda$ is the wavelength of the light 
-and $\theta_D$ the angle of the first order of diffraction.
-The intensity of the diffraction orders depends on the size and the response of one pixel.
-We see that we can decouple the effect of the periodicity of the grating, 
-that acts on the angles of the orders of diffraction, 
-and the response of one pixel, 
-that acts on the envelope of the angular response.
+This translates to an effective active pixel size of $d' < d$.\\
+
+In general, a grating gives rise to various diffraction orders
+with differing intensities
+and angles $\theta_p$, as dictated by the grating equation:
+$
+  \text{sin}(\theta_p) = p\lambda/d = p \, \text{sin}(\theta_D),
+$
+where $\lambda$ is the wavelength of the light,
+and $\theta_D$ is the angle of the first-order diffraction.
+The intensity of the individual diffraction orders is influenced by both the size and the response of a single pixel.
+Importantly, we can decouple the effects of the grating's periodicity,
+which influences the angles of the diffraction orders,
+from the effects of the response of a single pixel,
+which shapes the envelope of the angular response.\\
 
@@
   \includegraphics[width = \textwidth]{images/grating_geom.pdf}
-  \label{fig:flat_grating}
-
   \caption{
     \textbf{1D grating geometry.}
- ...
+    Schematic representation of the geometry of two types of modulators:
+    (a) the liquid crystal modulator, equivalent to a flat grating,
+    and (b) the DMD geometry, equivalent to a blazed grating.
+    $\alpha$ denotes the incident angle relative to the normal of the array plane,
+    $\theta_0$ refers to the angle of the zeroth diffraction order,
+    and $\theta_B$ is the tilt angle of the mirrors.\\
   }
@@
 
-For a normal incidence 
-in the case of a LC-SLM, the response of one pixel is 
-assumed constant over the pixel size. 
-For a blazed grating, we also have a linear phase slope on each pixel due to the tilt $\theta_B$ of the pixels.
-For an an arbitrary incidence angle $\alpha$, we also have a global phase slope 
-that has the effect of trivially shifting the angular diffraction by an angle $\alpha$.\\
-
-In essence, the incident angle $\alpha$ has the effect to shift the whole angular diffraction pattern 
-with respect to the normal incidence case, 
-while the blazed angle $\theta_B$, the tilt angle of the mirror in the DMD, 
-has the effect to shift only the envelope by $2\theta_B$.
-When the filling fraction is close to 100\%, 
-i.e. $d\approx d'$,
-the enveloppe for a flat grating is maximal 
-at $\theta_0=-\alpha$ corresponding to the angle of the zero-th order,
-and close to zero for the other orders.
-We then observe a single diffraction order, 
-which corresponds to the ideal case.
-Adding a blazed angle $\theta_B$ leads to a shift of the envelope, 
-the position of the maximum 
-$\theta_0=2\theta_B-\alpha$ 
-may not correspond anymore to a diffraction order.\\
+For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
+we can assume that the response of a single pixel is uniform its surface.
+Similarly, in the scenario of a blazed grating,
+such as a DMD, a linear phase slope is present on each pixel.
+This is due to the tilt angle $\theta_B$ of the mirrors.
+For an arbitrarily selected incidence angle $\alpha$,
+a global phase slope is introduced.
+This results in a trivial shift of the angular diffraction pattern by an angle $\alpha$.
+In essence, the incident angle $\alpha$ serves to shift the entire angular diffraction pattern
+relative to the case of normal incidence,
+while the blazed angle $\theta_B$—the tilt angle of the mirror in the DMD—
+only shifts the envelope by an angle of $2\theta_B$.
+Whenever the fill factor approaches 100\%,
+i.e. when $d \approx d'$,
+the envelope for a flat grating achieves its maximum at $\theta_0 = -\alpha$;
+this corresponds to the angle of the zeroth-order diffraction,
+which nears zero for all other orders.
+In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
+The addition of a blazed angle $\theta_B$ results in both
+a shift in the envelope and in the position of its maximum,
+now indicated by $\theta_0 = 2\theta_B - \alpha$.
+In the general case,
+this position may not align with a diffraction order anymore~\cite{Park2015properties}.
 
 \begin{optional}
-  If we neglect the finite size of the device and illumination 
-  and place ourselves in the small angle approximation, 
-  we can write the field reflected off the device 
-  for a plane wave illumination
-  in the two systems as:
+  Assuming the effect of the device's finite size and illumination to be negligible,
+  and situating ourselves within the context of the small-angle approximation,
+  we can represent the field reflected from the device
+  under the influence of plane wave illumination
+  across two systems as follows:
 
@@
     \begin{aligned}
-    R_\text{flat}(x) &\propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
-    R_\text{blazed}(x) &\propto 
+      R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
+      R_\text{blazed}(x) & \propto
       \underbrace{
@@
   by the absolute value squared of
-  Fourier transform of Eq.~\ref{eq:grating_response}
-  and reads:
+
+  The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
 
@@
       I_\text{flat}(\theta) \propto
-        \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D) 
+      \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
       \times
-        \text{sinc}^2\left( (\text{sin}(\theta)+\text{sin}(\alpha)) \frac{\lambda}{d'}\right)\\
+      \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
       I_\text{blazed}(\theta) \propto
@@
       \underbrace{
-          \text{sinc}^2\left( (\text{sin}(\theta)+\text{sin}(2\theta_B-\alpha)) \frac{\lambda}{d'}\right)
+        \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
       }_\text{envelope} \, .
@@
 and other orders also have a non-negligible contributions.
-% This implies that by selecting one order, we can loose more that 50\% of the light. 
-% In a 2D system, we can then loose more that 75\% of the light.\\ 
-The ideal case, i.e. when the maximum of the envelope corresponds to a diffraction order,
-leading a single diffraction order containing most of the energy,
-is obtained when the blazed grating equation is satistied~\cite{Casini2014on}:
+In the optimal scenario, where the peak of the envelope corresponds to a diffraction order,
+it results in a single diffraction order carrying the majority of the energy.
+This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}:
 
@@
 % \begin{equation}
-%     a = 
-%   \left\{
-%     \begin{aligned}
-%         a\\b
-% %         x + y   &= 7 \\
-% %         5x - 2y &= -7
-%     \end{aligned}
-%   \right\}
-% \end{equation}
-
-% \begin{equation}
 %   R_\text{flat}(x) \propto \Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d) 
@@
 To analyze more precisely the effect of diffraction in a DMD,
-we need to consider the 2D surface of the modulator.
-Let's define a cartesian coordinate system in the plane of the DMD
-with axis $x$ and $y$ aligned with the pixel sides 
-(see Fig.~\ref{fig:2d_geom}a.). 
-Pixels are repeated along those directions. 
-A technical difficulty is that the axis of rotation of the pixels 
-is the diagonal of the pixels, 
-hence rotated by 45 degrees with respect to the $x$ and $y$. 
-For the alignment and the maniputlation of the optical setup
-it is more convenient to work in incident and outgoing beams 
-with optical axis contained in an horizontal plane. 
-A simple and common solution consists in roatating the chip by 45 degrees
-with respect to the horizontal plane 
-such that the axis of rotation of the pixels is vertical.  
-This configuration is illustrated in Fig.~\ref{fig:2d_geom}.
+one needs to consider the 2D surface of the modulator.
+We can establish a Cartesian coordinate system on the plane of the DMD,
+with axes $x$ and $y$ aligned with the pixel sides
+(refer to Fig.~\ref{fig:2d_geom}a).
+Pixels are uniformly repeated along these axes.
+However, a technical challenge arises
+in that the axis of rotation of the pixels
+aligns with the pixel diagonals,
+resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
+For the convenience of alignment and manipulation of the optical setup,
+it is preferable to work with the incident and outgoing beams
+which have the optical axis contained in the horizontal plane.
+A straightforward and prevalent solution is to rotate the chip by 45 degrees
+relative to the horizontal plane,
+which aligns the pixel’s axis of rotation to be vertical.
+This configuration is depicted in Fig.~\ref{fig:2d_geom}b.
 
@@
   \centering
-  \includegraphics[width = 0.8\textwidth]{images/dmd_45.pdf}
+  \includegraphics[width = 0.9\textwidth]{images/dmd_45.pdf}
   \caption{
     \textbf{2D grating geometry.}
- ...
+    (a) Schematic representation of the geometry of the DMD.
+    (b) Photograph of the DMD chip oriented so that the rotation axis of the pixels
+    is aligned vertically.
   }
@@
 
-The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions.
-The incident angle $\alpha'$. 
-Because of the 45 degrees angle between 
-axis of rotation of the mirrors and axis, 
-both 1D gratings have the same properties. 
-A more complete computation of the 2D system can be found in~\cite{Scholes2019structured}.
-To apply Eq.~\ref{eq:blazed_eq}, 
-we first need to project the 
-the different angles of the problem in the incident planes 
+The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions,
+both having the same properties.
+A more comprehensive depiction of the 2D system can be found in~\cite{Scholes2019structured}.
+In order to utilize Eq.(\ref{eq:blazed_eq}),
+one first needs projecting
+the different angles of the problem onto the incident planes
 of the two 1D gratings.
-It gives 
+This yields
 $\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$,
@@
 and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$,
-$theta_\text{DMD}$ being the actual rotation angle of the mirrors with respect to the diagonal of the pixels.
+where $\theta_\text{DMD}$ represents the true rotation angle of the mirrors relative to the diagonal of the pixels,
+and $\alpha'$ and $\beta'$ denote the incident and outgoing angles in the horizontal plane.
 We can quantify how close we are to the ideal case,
-i.e. when satifying the blazed equation~\ref{eq:blazed_eq}, 
-by defining a {\em blazed number} $\mu$ as~\cite{WFSnet_diffraction}:
+i.e., when satifying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
+by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:
 
@@
   \right| \, ,
+  \label{eq:blazed_number}
 \end{equation}
@@
 and $\mod{2}$ the modulo 2 operation.
-$\mu$ is maximal when the blazing equation is satisfield, 
+$\mu$ is maximal and equals $1$ when the blazing equation is satisfield,
 i.e. when one order of diffraction contains most of the energy,
@@
 
-To illustrate the effect, we simulate a DMD in Python 
-(see tutorial and code in~\cite{WFSnet_diffraction})
-with a two pixel pitches of 
+To demonstrate the effect, we conduct a simulation of a DMD using Python
+(refer to tutorial and code in~\cite{WFSnet_diffraction})
+with two pixel pitches of
 $d=7.6$\textmu m and $d=10.8$\textmu m,
-for a coherent excitation at $\lambda=633$nm. 
-We show in Fig.~\ref{fig:mu76} the blazed number $\mu$
-as a function of the incident angle $\alpha'$ 
-as well as the far field diffraction pattern for different two incident angles.
-While one can change the efficiency of diffraction by tuning 
-the incindent angle, its effect is somewhat limited 
-withing a reasonable angular range compatible with experimental constraints 
+under a coherent excitation at $\lambda=633$nm.
+Fig.~\ref{fig:mu76} shows the estimated blazed number $\mu$
+as a function of the angle of incidence $\alpha'$,
+along with the far field diffraction pattern for two distinct incident angles.
+It should be noted that the efficiency of diffraction can be altered by adjusting
+the angle of incidence, yet its impact is relatively confined
+within an acceptable angular range that aligns with experimental limitations
 (i.e. for angles far from $\pm 90^\circ$).
-Far field pattern are centered around the maximum of the envelope 
+Far field patterns are centered around the maximum of the envelope
 $\theta_\text{max} = 2\theta_B - \alpha$
@@
   \caption{
-  \textbf{Blazed number and far field diffraction pattern for...}
- ...
+    \textbf{Blazed number and far field diffraction patterns, extreme cases.}
+    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha'$  (top)
+    for a pixel pitch of $d=7.6$\textmu m (left)
+    and $d=10.8$\textmu m (right).
+    Corrseponding far field diffraction pattern (bottom)
+    for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
+    The yellow cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
   }
@@
 Having a small value for the blaze number $\mu$
-does not simply limit the amount of light modulated 
-due to the reduced diffraction efficiency, 
-it also impact the modulation qualoty by introducing cross-talk 
+not only restricts the amount of light modulation
+due to the diminished diffraction efficiency,
+it also influences the modulation quality by inducing cross-talk
 between the two states of the DMD pixels.
-For now, we considered that all the pixels are in the same state. 
-In a practical use of the DMD,
-we need to modulate the state of each pixel individually.
-When $\mu$ is close to zero, higher order of diffraction 
-still have a non negligible intensity, 
-as shown in the 1D case in Fig~\ref{fig:gratings}. 
-One consequence is that pixels in the {\em off} state 
-may have orders of diffraction that are not blocked by the pinhole, 
-and will contribute as a perturbation to the modulated wavefront. 
-While this contrinution may seem small, 
-it does impact the quality of the modulation 
-as the modulation scheme usually requires about half the pixels 
-to be in the {\em off}  state for phase modulation~\cite{}, 
-and even more for complex modulation schemes~\cite{}.\\
+Until now, we have assumed that all the pixels are in the same state.
+In actual usage of the DMD,
+it becomes necessary to modulate the state of each pixel individually.
+When $\mu$ approaches zero, higher orders of diffraction
+still possess a significant intensity,
+as demonstrated in the 1D case in Fig~\ref{fig:gratings}.
+One adverse implication is that pixels in the {\em off} state
+may contain orders of diffraction that are not blocked by the pinhole,
+and therefore, will contribute as an interference to the modulated wavefront.
+While this contribution might appear weak,
+it does affect the quality of the modulation
+since the modulation scheme typically necessitates about half the pixels
+to be in the {\em off} state for phase modulation~\cite{lee1979binary},
+and even more so for elaborate modulation schemes~\cite{goorden2014superpixel,Gutierrez2024DMD}.\\
 
@@
     \textbf{Cross-talk between {\em on} and {\em off} states.}
- ...
+    We show the computed normalized amplitude of the diffraction patterns corresponding
+    to all the pixels in the {\em on} state (blue)
+    and {\em off} state (orange)
+    for two different pixel pitches,
+    $d = 7.6$\textmu m (a.)
+    and $d = 10.8$\textmu m (b.)
+    with the same experimental conditions.
+    In the first case, $\mu$ is close to $0$,
+    we observe a non-negligible contribution of the {\em off} state
+    at the main diffraction order of the {\em on} state,
+    thus creating unwanted cross-talk.
+    In the second case, $\mu$ is close to $1$,
+    the contribution of the {\em off} state is negligible.
   }
@@
   \textbf{TL;DR:}\\
-  DMDs act as a blazed grating. 
+  DMDs act as a blazed gratings.
   For a given operation wavelength,
@@
   and diffraction efficiency.
+  It can be done by estimating the \textit{blazed number} $\mu$ introduced in eq.~(\ref{eq:blazed_number})
+  or directly using our custom online tool~\cite{popoffDMDDiffractionTool}.
 \end{tldr}
@@
 
-While only able to provide a hardware binary amplitude modulation,
-DMDs are a powerful tool for wavefront shaping and sensing. 
-For these applications, 
-aberrations due to the non-flatness of the DMD surface 
-are detrimental and need to be characterized and corrected for. 
-For LC-SLMs, inhomoegeneities of the surface is usually characterized by the 
-manufacturer and the spatial phase profile of the introduced aberrations 
-in the plane of the modulator is provided for the working wavelength.
-As previously mentioned,
-when used for intensity modulation in a plane conjugated to the DMD plane, 
-as used in digital projectors, 
-the system is not sensitive to aberrations brought by the DMD surface.
-For these reasons, this effect is usually not documented, 
-and usually not mentioned, in the information provided by the manufacturer 
-despite the fact that the deviation from a flat surface
-typically reaches few wavelengths~\cite{Brown2021multicolor}.
-\\
+While only capable of providing hardware binary amplitude modulation,
+DMDs serve as an potent tool for wavefront shaping and sensing.
+These applications  critically require characterization and correction of aberrations
+caused by the non-flatness of the DMD surface.
+For LC-SLMs, the manufacturer typically characterizes the surface's inhomogeneities within the plane of the modulator and
+provides a spatial phase profile of the introduced aberrations based on the operational wavelength.
+It is noteworthy that when utilized for intensity modulation on a plane conjugated to the DMD plane,
+as it is the done in digital projectors,
+the system becomes insensitive to the aberrations caused by the DMD surface.
+Consequently, these effects are commonly overlooked and rarely documented in the information provided by the manufacturers.
+They are nonetheless non-negligible, with deviations from a flat surface
+that typically spans across several wavelengths~\cite{Brown2021multicolor}.\\
 
@@
 
-In the litterature, different methods have been developped to characterize 
-the phase pattern representing the aberrations of the DMD in the 
-plane of the modulator. 
-Typically, it consists in using a model of the aberations 
-and adjusting the parameters to fit the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
-While providing accurate results, 
-the implementation of the approach has to be adapted to the specific 
-setup condition and may require
-additional optical components, 
-precise alignments, 
+In the literature, various methods have been proposed to characterize
+the phase pattern of the DMD aberrations in the plane of the modulator.
+Typically, this involves the use of a model for the aberrations,
+tweaking the parameters to align with the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
+An alternative approach entails direct measurement of the distorted wavefront,
+either employing a wavefront sensor such as a Shack-Hartmann~\cite{Lee2023compensation},
+or via interferometry.
+While such methods yield accurate results,
+they necessitate adaptation to the particular
+setup conditions,
+and frequently require supplementary optical components,
+meticulous alignment,
 and custom software.
-An other approach consists directly measuring the distorded wavefront, 
-either using wavefront sensor, such as a Shack-Hartmann, 
-or using interferometry. 
-Again, it requires specific hardware and 
-a carefully calibrated and stable setup.
-In this section, we present a simple method to characterize the aberrations using a lens and a camera. 
-This approach can be used for any system prividing a phase modulution, 
-such as LC-SLMs or deformable mirrors.\\
+Furthermore, they also demand
+a meticulously calibrated and stable setup.
+In this section, we introduce a straightforward method to characterize the aberrations using a lens and a camera.
+This technique can be employed for any system that offers phase modulation,
+such as LC-SLMs or deformable mirrors. \\
 
@@
     \textbf{Setup for aberration correction.}
- ...
+    (a) A schematic representation of the DMD array divided into macropixels,
+    of which the phase can be controlled.
+    The precise shape of the macropixels and the area controlled are not crucial for the characterization of aberrations to be effective,
+    provided that the spatial sampling is sufficient to capture the highest spatial frequencies of the aberrations.
+    (b) A schematic representation of the setup employed to characterize the aberrations.
+    A camera is used to image the far field on the DMD surface using a lens.
+    (c) shows the ideal intensity pattern corresponding to the numerical aperture of the illumination setup,
+    i.e. its theoretical ideal point spread function.
+    (d) show the actual recorded intensity pattern corresponding to a distorted point spread function due to the DMD aberrations.
   }
@@
 
-He assume that the DMD is configured to provide a phase modulation~\cite{Lee1979binary, RODRIGO}. 
-It implies that the modulator can be devided into regions, we call {\em macropixels},
-on which the phase can be independently controlled.
-We place a lens and a camera in its Fourier plane and illuminate the modulator with a collimated beam 
-covering the full desired area of the modualtor we intend to use. 
-If there is no or negligible aberrations, the observed intensity pattern 
-reseembles the PSF of the lens, i.e. an Airy disk such as presented in Fig.~\ref{fig:dmd_aberr_setup}.c.
-However, in partice, we observe a significantly distorted pattern,
-such as the one presented in Fig.~\ref{fig:dmd_aberr_setup}.d.\\
-
-We assume the the aberrations introduced by the DMD is smooth and 
-can be described by a phase pattern $\phi^\text{aber}$ 
-that can be well approximated by a finite number of 
-Zernike polynomials $Z_n(r,\theta)$~\cite{Zernike1934beugungstheorie}, such as:
+We assume that the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
+This implies that the modulator can be divided into $N$ sections,
+which we designate as {\em macropixels},
+where the phase can be controlled independently.
+We use a lens and a camera in its Fourier plane,
+illuminating the modulator with a collimated beam
+that extends over the entire area of the modulator intended for use.
+In scenarios where there are minimal or no aberrations,
+the intensity pattern observed would mimic the PSF of the lens,
+such as an Airy disk depicted in Fig.~\ref{fig:dmd_aberr_setup}.c.
+However, in practice, we encounter a substantially distorted pattern,
+like the one represented in Fig.~\ref{fig:dmd_aberr_setup}.d.\\
+
+We hypothesize that the aberrations brought about by the DMD are smooth,
+and can be depicted by a phase pattern $\phi^\text{aber}$ in the plane of the DMD array.
+This could be feasibly approximated by a finite number of Zernike polynomials
+$Z_n(r,\theta)$~\cite{Zernike1934beugungstheorie}, as follows:
 
 \begin{equation}
-  \phi^\text{aber}(r,\theta) \approx \sum_{n=0}^N a_n Z_n(r,\theta) \, .
+  \phi^\text{aberr}(r,\theta) \approx \sum_{n=0}^N a_n Z_n(r,\theta) \, .
 \end{equation}
 
-The goal is to find and display the phase pattern 
-on $\phi_i^\text{corr}$ on each pixel $i$ that best compensate for the aberrations, 
+The goal is to find and display the phase value
+$\phi_i^\text{corr}$ on each macropixel $i$ that best compensates for the aberrations,
 i.e. $\phi_i^\text{corr} = -\phi^\text{aber}(r_i,\theta_i)$.
@@
     \textbf{Correction phase map.}
- ...
+    (a) Schematic representation of the function we want to optimize.
+    We both want to maximize the intensity of the main lobe of the point spread function
+    and minimize its side lobes.
+    (b) Values of the experimentally obtained optimal coefficients in the Zernike polynomials basis.
+    (c) Resulting optimal phase pattern in the chosen illumination area.
   }
@@
 
-We will perform a sequential optimization of the parameters $a_n$ 
-to maximize a cost function that is optimial for an ideal correction of the abberations.
-To define such a cost function, we first create a mask with a disk 
-centered around the maximum of intensity of the initial image (resembling Fig.~\ref{fig:dmd_aberr_setup}.d)
-and of radius equivalent to a speckle grain. 
-This size does not need to be very very precisely set for the optimization procedure to work, 
-and can be found by estimating the size of the ideal PSF, i.e. $r_0 \approx M \frac{\lambda}{2 NA}$, 
-with $NA$ the numerical aperture of the optical system and $M$ its magnification.
-For a given output image,  we compute the element-wise product between the intensity image measured and the mask and perform a summation.
-We start with $a'_n = 0\,\, \forall n \in [0..N]$.
-For each parameter, we try different values of $a'_n$,
-generate the phases on each pixel so that 
+We perform a sequential optimization of parameters $a_n$
+to maximize a specific function designed
+to be maximal for the ideal correction of optical aberrations.
+We first  generate a mask, represented by a disk,
+centered around the point of maximum intensity of the original image (refer to Fig.~\ref{fig:dmd_aberr_setup}.d)
+with a radius equivalent to a single speckle grain.
+The exact size of this radius for a successful optimization is not critical
+and can be determined by approximating the dimensions of the ideal point spread function,
+expressed as $r_0 \approx M \frac{\lambda}{2 NA}$,
+where $NA$ is the numerical aperture of the optical system and $M$ refers to its magnification.
+For a given output intensity pattern, we compute an element-wise product between this image
+and the created mask, followed by a summation.
+This calculated sum is then divided by an analogous product, but with the complementary mask substituted in place of the original
+in order to minimize the side lobes of the point spread function.
+We set initial parameter values as $a'_n = 0\,\, \forall n \in [0..N]$.
+For each parameter, we test different values of $a'_n$,
+construct the phases for every micropixel according to
 $\phi_i^\text{corr} = \sum_{n=0}^N a'_n Z_n(r_i,\theta_i)$,
-measure the output intensity profile,
-and estimate the value the cost function.
-We keep for each parameter the value that leads to the highest value.
-We repeat the whole process $3$ times for each parameter to 
-mitigate the effect of noise or vibration.
+record the resulting intensity profile,
+and evaluate the corresponding cost function.
+For each parameter, the value that results in the maximum output is retained.
+To mitigate potential noise or instability, this complete process is reiterated $3$ times for each parameter.\\
+
+To estimate the quality of the correction,
+we compute the Strehl ratio of the point spread function.
+It is defined as the maximum of the measured PSF divided
+by the maximum of the ideal one.
+This optimal PSF is
+the  squared modulus of the Fourier transform of a circular aperture~\cite{airy1835diffraction}:
+
+\be
+PSF_\text{ideal} \propto
+\left[
+  J_1\left(k a \frac{R}{\sqrt{R^2+f^2}}\right)
+  \right]^2 \, ,
+\ee
+
+with $J_1$ the Bessel function of the first kind of order $1$,
+$k = 2\pi/\lambda$ the wavenumber,
+$a$ the radius of the aperture,
+$R$ the radial coordinate in the Fourier plane,
+and $f$ the focal length of the lens.\\
+
+As an illustration, we conduct an optimization procedure using 11 Zernike polynomials.
+We use a V-9501 Vialux DMD with a DLP9500 TI chip
+of resolution 1920 by 1200 pixels and a pixel pitch of 10.8 \textmu m.
+The optimization is performed on a disk of radius $340$ pixels,
+The illumination is done using an expanded laser beam at 633 nm,
+corresponding to the aperture of our optical setup.
+We exclude the first three Zernike polynomials in the optimization process,
+starting from the radial degree $2$.
+Indeed, the initial one, known as the piston, has no influence on the PSF quality.
+The subsequent ones, the tip and the tilt, cause the PSF to shift.
+Our procedure relies on optimizing the maximum of the PSF, wherever that is,
+rendering us indifferent to these two parameters.
+After optimization,
+it is possible to generate the correction pattern using a selected number of Zernike polynomials
+in order to investigate their impact on the PSF quality.
+We see here that using about 10 Zernike polynomials is sufficient to obtain a Strehl ratio $>0.99$.
+Fig.~\ref{fig:zernike} demonstrates the Strehl ratio and the intensity profiles of the PSF
+for different counts of the utilized Zernike polynomials.
+It is important to note that we do not use the full surface of the DMD.
+Using a larger area may lead to stronger deformations of the PSF,
+requiring a larger number of Zernike polynomials to be corrected accurately.
+The experimental data, in addition to the Python code used to generate the figures, can be accessed in the dedicated repository~\cite{github}.\\
 
@@
     \textbf{Effect of Zernike polynomials on the optimizaton of the PSF.}
- ...
+    (a) Strehl ratio of the point spread function as a function of the number of Zernike polynomials used
+    for the compensation of the aberrations.
+    (b) Intensity profile of the point spread function for different number of Zernike polynomials
+    in linear and logarithmic scale.
+    To generate this data, we take the final solution of the optimization experimental procedure,
+    i.e. the optimal coefficients $a'_n$,
+    and generate the corrected PSF with an increasing number of Zernike polynomials.
   }
@@
 
-We provide in the section Python pieces of code to perform the aberration correction.
-We assume that the phase modulation can be controlled using a function 
-!display_phase_mask(phase_pattern)! to display a desired phase pattern 
-containing real values between $2\pi$.
-We remind that phase modulation using a DMD requires an appropirate setup 
-and procedure for the generation of the binary pattern that is not detailed here 
-but detailed in ths tutorial~\cite{RODRIGO}.
-We also assume a function !get_camera_image()! 
-that returns the intensity pattern measured by the camera. 
-We use the package !aotools!~\cite{Townson2019aotools} for generating Zernike polynomials.
-
-% \lstinputlisting[language=Python, caption=Python example]{code/dmd_aberrations.py}
-
-measuring projection of a signal wavefront with a displayed mask and a bucket detector~\cite{}
-
-\cite{Lee1979binary}
+In this section, we provide Python code snippets for performing aberration correction.
+We presuppose that phase modulation can be controlled using a function,
+!display_phase_mask(phase_pattern)!, to display a desired phase pattern
+with real values between $0$ and $2\pi$.
+It is important to note that phase modulation utilizing a DMD demands an appropriate setup
+and process. Although not detailed here,
+this is elaborated upon in the referenced tutorial~\cite{Gutierrez2024DMD}.
+Additionally, we assume the existence of a function, !get_camera_image()!,
+which returns the intensity pattern as measured by the camera.
+For generating Zernike polynomials, we make use of the !aotools! package~\cite{Townson2019aotools}.
+
+\begin{tldr}
+  \textbf{TL;DR:}\\
+  DMDs are not flat and can introduce aberrations;
+  these are typically much stronger than those commonly observed with liquid crystal SLMs.
+  This can be counteracted in situ using a standard setup and a straightforward optimization procedure
+  to maximize the intensity at the central position of the point spread function.
+\end{tldr}
 
@@
 
-Most DMD kits are composed of two elements, 
-the chip itself and the electronics board to control it. 
-It can be the standard electronics board for video projectors,
-as it is the case for TI evluations kits, 
-or a FPGA for fast scientific use, 
-as provided by Vialux~\cite{vialux}. 
-These electronics come with a fan to cool the chip and the electronics board. 
-As the connection between the chip and the electronics board 
-used a rigid flat cable, 
-the two elements are not mechanically connected.
-Vibrations from the board are then transmitted to the chip,
-leading to small angle variations of the mirror surface. 
-While this effect is without consequences for video projection, 
-it can be dramatic applications to complex media 
-as their response is highly sensitive to the phase front.\\
-
-Because of this high sensitivy, and since we want to use the DMD 
-in the context of complex media, it is simpler to characterize this effect 
-directly on the system response rather than 
-building a different setup to characterize the wavefront itself.
-We show an example of such setup in Fig.~\ref{fig:MMF_ref}. 
-We expand an laser beam onto the DMD and sent the input light through a multimode fiber. 
-While we typically want phase modulation, 
-the setup will need additional elements to achieve this task~\cite{}. 
-We here simplify the setup for the sake of clarity. 
-The output of the fiber interfere with a reference arm in order to measure 
-changes of the field.
-We then measure the interference pattern using a camera. \\
-
-\begin{figure}
+Most DMD kits consist of two primary components,
+the chip itself and the electronic board that controls it.
+This could be the standard electronics board typically used for video projectors,
+as seen in TI evaluation kits,
+or an FPGA specifically designed for rapid scientific usage,
+as offered by Vialux~\cite{vialux} for instance.
+Integral to these electronics is a fan designed to cool both the chip and the electronic board.
+However, due to the use of a rigid flat cable for connection between the chip and the electronics board,
+these parts are not mechanically independent.
+As such, vibrations originating from the board are partially transmitted to the chip,
+resulting in minute rotations of the mirror surface.
+Although this perturbation is inconsequential for video projection,
+they can have significant impacts on applications involving complex media
+given their high sensitivity to phase front variations.\\
+
+Due to its high sensitivity in complex media,
+it is convenient to characterize this effect
+directly on the system's response,
+rather than constructing a distinct setup to analyze the wavefront itself.
+An example of such a setup is demonstrated in Fig.~\ref{fig:MMF_ref},
+although a similar approach can be employed with a scattering medium.
+We enlarge a laser beam onto the DMD and transmit the incoming light through a multimode fiber.
+Additional elements are required in the setup to fulfill the requirement for complex modulation~\cite{Gutierrez2024DMD}.
+For the sake of clarity, we present a simplified version of the setup where those elements are not present.
+The output from the fiber is then made to interfere with a reference arm
+in an off-axis configuration~\cite{Cuche2000spatial},
+allowing us to detect changes in the output complex field
+by recording the interference pattern using a camera.\\
+
+\begin{figure}[ht]
   \centering
@@
     \textbf{Measuring stability at the output of a multimode fiber.}
- ...
+    Schematic representation of the setup used to measure the phase fluctuations
+    through a multimode fiber.
+    An equivalent setup can be used with a scattering medium.
   }
@@
 
-We show in the supplementar maeterials~\cite{} an animation of the dynamic pattern. 
-We also estimate the phase variation as a function of time. 
-In a given region of the image, 
-the relation between the phase and the transverse displacement of the fringes
-is proportional to the phase, 
-with a displacement equal to the fringes period corresponding to $2\pi$.
-We show in Fig.~\ref{fig:phase_vibrations} (red curve) 
-the fluctuation of the phase over time presenting fast oscillations
-due to the mechanical vribration communicated by the board.\\
-
-\begin{figure}
+In the supplementary materials~\cite{}, we present an animation illustrating the dynamic pattern.
+In off-axis holography,
+the local transverse displacement of the fringes
+is directly proportional to the phase,
+with a displacement equivalent to the period of the fringes corresponding to $2\pi$~\cite{Cuche2000spatial}.
+This permits us to estimate the fluctuation in phase over time at a given postition of the output plane.
+As illustrated in Fig.~\ref{fig:phase_vibrations} (depicted by the red curve),
+the phase varies rapidly over time,
+a fluctuation attributed to the rapid mechanical vibrations transmitted by the board.\\
+
+\begin{figure}[ht]
   \centering
@@
   \caption{
-  \textbf{Vibration indiced phase fluctuations.}
- ...
+    \textbf{Vibration induced phase fluctuations.}
+    Phase fluctuations measured over time at
+    a specific position on a plane located at the distal end of a multimode fiber.
+    We employ off-axis holography using the setup depicted in Fig.~\ref{fig:MMF_ref}.
+    The red and blue curves correspond to the phase fluctuations
+    measured respectively without and with vibration damping
+    by securing the flat cable with foam, as illustrated in Fig.~\ref{fig:dumping}.
   }
@@
 
-A simple but efficient solution consists in dumping the vibrations
-at the level of the flat cable by clamping the cable with soft maeterial, 
-as illustrated in Fig.~\ref{fig:dumping}. 
-It can be done with off the shelf material.
-We use here simple foam material used for packaging clamped to the cable 
-using two metallic plates, screws, and nuts.
-We observe a significant reduction of the phase fluctuations 
-as shown in Fig.~\ref{fig:phase_vibrations} (blue curve).\\
-
-\begin{figure}
+An simple yet effective solution consists in dampening the vibrations
+at the flat cable's level by clamping it with a soft material,
+as depicted in Fig.~\ref{fig:dumping}.
+This can be achieved using commonly available materials.
+In this context, we utilize simple foam, typically used for packaging, and secure it to the cable
+with two metallic plates, screws, and nuts.
+We observe a significant decrease in the phase fluctuations,
+as demonstrated in Fig.~\ref{fig:phase_vibrations} (blue curve).\\
+
+\begin{figure}[ht]
   \centering
@@
     ...
-  }
     \label{fig:dumping}
-\end{figure}
+  }
+  \label{fig:dumping}
+\end{figure}
+
+\begin{tldr}
+  \textbf{TL;DR:}\\
+  The functioning of DMDs can be perturbed by vibrations transmitted from the electronics board
+  via the rigid flat cable that links it to the DMD chip.
+  This adverse effect can be minimized by securing the cable with a soft material
+  that serves to dampen these vibrations.
+\end{tldr}
 
@@
 
-Electronics used to control the DMD chip heat up during operation.
-This effect depends on the frame rate. 
-In particular, it is important to notice that this effect is important when 
-a sequence is running on the DMD, and is limited when the DMD is powered 
-but without any sequence running.
-Changes of temperature has the effect of deforming the surface of the chip 
-and potentially changing the phase introduced by the glass protecting window.
-This has the effect of introducing low order aberrations that degrade the quality of the wavefront.
-While this effect is relatively less important than the ones due to 
-static aberrations and mechanical instabilities detailed previously, 
-it still have a significant impact on the response of a complex medium 
-when using a DMD to modulate the input wavefront.\\
-
-The first step consists in characterizing the effect of temperature. 
-While a precise effect on the wavefront itself can be measured~\cite{Rudolf2021thermal}, 
-it is once again usually more convenient and relevant to measure the effect directly on the response of the system studied.
-
-Thermal stability~\cite{Rudolf2021thermal}
+Electronics utilized to control the DMD chip experience thermal variations during operation.
+The dynamics of this effect are dependent on the frame rate.
+Specifically, the chips heats up more quickly when increasing the frame rate.
+The increase of temperature can reach more that $15$ degrees Celsius
+when running a sequence at maximum speed (20-30 kHz)~\cite{Rudolf2021thermal}.
+Notably, this effect is less pronounced when the device is on but not running a sequence.
+Temperature fluctuations can cause deformations on the chip's surface
+and can modificy the phase response of the glass protective window.
+This creates low order aberrations, which degrade the quality of the wavefront.
+While this issue is comparatively less critical than static aberrations and mechanical instabilities
+previously detailed,
+it nonetheless has a substantial impact on the complex medium's response
+when a DMD is employed to modulate the input wavefront.\\
+
+Before initiating a wavefront shaping experiment,
+it is important to characterize the influence of temperature
+to assess the extent to which it impacts the results.
+Although the exact effect on the wavefront distortion can be directly quantified~\cite{Rudolf2021thermal},
+it is typically more convenient
+to directly measure the effect on the studied system's response.
+To do so, we use a setup to the one presented in the previous section
+and depicted in Fig.~\ref{fig:MMF_ref}.
+We can then estimate the field or intensity decorrelation over time.
+We present here results with intesity correlation, as it does not require a reference arm.
+The measured output pattern typically take the form of a seemingly random speckle pattern,
+that is sensitive to minute changes of the input wavefront.
+The correlation estimation is obtained by comparing the output intensity pattern
+$I(\vec{r}, t)$ at a given time $t$
+to the one at $t=0$.
+We use the following expression for the correlation:
 
@@
     }
-  }
+  } \, ,
+  \label{eq:temp_decorr}
 \end{equation}
 
+with $\bar{I}(\vec{r}, t) = I(\vec{r}, t) - \left\langle I(\vec{r}, t) \right\rangle_{t}$
+and $\left\langle . \right\rangle_{\vec{r}}$.
+$\left\langle . \right\rangle_{t}$
+represent the spatial averaging over the region of interest
+of the output plane (i.e. the plane of the camera sensor)
+and the temporal averaging over the measured frames.
+Fig.~\ref{fig:temp_decorr} shows the measured decorrelation
+over time when running an sequence in an infinite loop.
+$t=0$ correspond in the first case (a)
+to the start of a sequence
+and is set in the second one to
+4.2 hours post the after of the sequence (b).
+For the scenario (a), it is to note that before the sequence commenced,
+the DMD was active but remained in the idle state,
+meaning that it was not executing a sequence.
+We observe a non-negligible decorrelation over time
+after the sequence started
+that slows down after about 1 hour.
+When running the same experiment 4.2 hours after starting the sequence,
+the correlation is now stable.\\
+
 \begin{figure}[ht]
@@
     \textbf{Effect of temperature.}
-    ...
-  }
-  \label{fig:Cizmar_1}
-\end{figure}
+    Temporal decorrelation of a speckle pattern measured at the output of a multimode fiber
+    when running a sequence in an infinite loop.
+    We use Eq.(\ref{eq:temp_decorr}) to estimate the correlation between the intensity pattern
+    at a given time $t$ and the one at $t=0$.
+    $t=0$ corresponds respectively to the start of the sequence (a)
+    and to 4.2 hours after the start of the sequence (b).
+  }
+  \label{fig:temp_decorr}
+\end{figure}
+
+The more efficient way to counteract this effect is to use a
+close-loop system to stabilize the temperature of the DMD chip.
+This can be achieved by using a thermoelectric cooler
+as demonstrated in Ref.~\cite{Rudolf2021thermal}
+and depicted in Fig.~\ref{fig:decorr_T}.
 
@@
     \textbf{Thermal stabilization of the DMD chip.}
+    Schematic representation of the setup utilized to regulate the temperature of the DMD chip.
+    A thermoelectric cooler and a heat sink is employed to control the temperature,
+    while a thermistor is used to measure deviations from the temperature setpoint.
+    A closed-loop system is used to regulate the temperature.
     Image adapted from~\cite{Rudolf2021thermal}.
@@
 
-\section{Control Vialux DMDs with Python}
-
-\cite{hofling2004alp}
-
-\cite{popoff2016alp4lib}
+% \section{Control Vialux DMDs with Python}
+
+% \cite{hofling2004alp}
+
+% \cite{popoff2016alp4lib}
 

### 2023-11-16 paper/main.tex
remove ALP4lib part
Add outline

@@
 and will not be elaborated upon here.
-For the remainder of this paper, it will be assumed that the DMD is used for complex modulation via such a method.
-While the present tutorial is not constrained to any specific device,
-the software section will predominantly focus on the Vialux control board.
-In the first part....\\
+For the remainder of this paper, it will be assumed that the DMD is used for complex modulation via such a method.\\
+% While the present tutorial is not constrained to any specific device,
+% the software section will predominantly focus on the Vialux control board.
+In the first part, we present the diffraction properties of a DMD
+and detail how they may affect the efficiency of the system.
+We provide a simple criteria for choosing the correct DMD parameters for a given excitation wavelenght.
+In the second part, we detail the aberration effects induced but the non-flatness of the DMD surface.
+We show how to simply characterize and compensate this effect.
+In the third section, we describe the effect of mechanmical vibrations induced by the cooling system of the DMD.
+Finally, we show how the thermalization of the DMD chip can induce fluctuations of the DMD response over time.
 

### 2023-11-20 paper/main.tex
remove \ee \be
test
Updates from Overleaf

--- 2023-11-15 paper/main.tex
@@
 \usepackage{subcaption}
+\usepackage{orcidlink}
 
@@
 
-\author{Sébastien M. Popoff}
+\author{Sébastien M. Popoff~\orcidlink{0000-0002-7199-9814}}
 \address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
@@
 
-\author{Rodrigo Gutiérrez-Cuevas}
+\author{Rodrigo Gutiérrez-Cuevas~\orcidlink{0000-0002-3451-6684}}
 \address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
@@
 
-\author{Yaron Bromberg}
+\author{Yaron Bromberg~\orcidlink{0000-0003-2565-7394}}
 \address{Racah Institute of Physics, The Hebrew University of Jerusalem, Israel}
@@
 Nonetheless, LC-SLMs are hampered by their slow response time,
-permitting only a modulation speed ranging from a few Hz to about 100 Hz.
+permitting only a modulation speed ranging from a few Hz to about 100 Hz.\\
+
 Digital Micromirror Devices (DMDs) have emerged as a technology bridging the gap
@@
 they offer a large number of pixels (similar to LC-SLMs) and fast modulation speeds (typically up to several tens of kHz).
-However, DMDs are restricted to binary amplitude modulation and are not optimized for coherent light applications.
+Their high speed capabilities made them attractive for real-time applications,
+in particular for high-resolution imaging microscopy
+requiring fast scanning or illumination shaping~\cite{Cha2000,Zhuang2020},
+biolithography~\cite{yoon2018emerging},
+and optical tweezers~\cite{Gauthier2016}.
+However, DMDs are restricted to hardware binary amplitude modulation and are not optimized for coherent light applications.
 Utilizing DMDs for coherent control of light in complex media is therefore non-trivial
-and necessitates specific adaptations for efficient use.
-While other articles exist describing the various aspects of DMDs~\cite{Park2015properties, Cox2021converting, Wang2023diffraction},
-this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
-In particular, we provide characterization and validation procedures that requires minimal changes compared to typical wavefront shaping setups.\\
-
-REVIEW en bio imaging~\cite{Zhuang2020}
+and necessitates specific adaptations for efficient use.\\
 
@@
 and its original design intentions.
-Investigated and developed by Texas Instruments since the late 1970s,
-DMDs gained prominence in the 1990s for video projection applications
-under the commercial name of Digital Light Processing (DLP)~\cite{Dudley2003emerging}.
+Investigated and developed by Texas Instruments since the 1980s,
+DMDs gained prominence in the 1990s for video projection applications since the 1990s
+under the commercial name of Digital Light Processing (DLP)~\cite{Hornbeck1997,Dudley2003emerging}.
 The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
@@
 and will not be elaborated upon here.
-For the remainder of this paper, it will be assumed that the DMD is used for complex modulation via such a method.
-While the present tutorial is not constrained to any specific device,
-the software section will predominantly focus on the Vialux control board.
-In the first part....\\
+For the remainder of this paper, it will be assumed that the DMD is used for complex modulation via such a method.\\
+% While the present tutorial is not constrained to any specific device,
+% the software section will predominantly focus on the Vialux control board.
+
+While other articles exist describing the various aspects of DMDs~\cite{Park2015properties, Scholes2019structured, Cox2021converting, Wang2023diffraction},
+this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
+In particular, we provide characterization and validation procedures that requires minimal changes compared to typical wavefront shaping setups.\\
+We first introduce the diffraction properties of a DMD
+and elaborate on how these could impact the system's efficiency.
+We also furnish a straightforward criterion for selecting the appropriate DMD parameters for a specified excitation wavelength.
+In the next section,
+we delve into the aberration impacts brought about by the non-flatness of the DMD surface.
+We demonstrate a simple process to characterize this effect and provide compensation solutions.
+In the third segment,
+we detail the influence of mechanical vibrations that are induced by the DMD's cooling system.
+Lastly, we discuss how the thermalization of the DMD chip can potentially result in variations to the DMD response over time.
 
@@
 
-\be
+\begin{equation}
   PSF_\text{ideal} \propto
@@
     \right]^2 \, ,
-\ee
+\end{equation}
 
@@
     \textbf{Dumping mechanical vibrations.}
-    ...
-    \label{fig:dumping}
+    Pictures of experimental setups where a damping of the vibrations is implemented.
+    (a) The flat cable is secured with foam and metallic plates.
+    (b) The flat cable is in tension against a foam material supported by a metallic plate.
   }

### 2023-11-20 paper/main.tex
correct git history

@@
 \address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
-\ead{rodrigo.gutierrez-cuevas@espci.psl.eu>}
+\ead{rodrigo.gutierrez-cuevas@espci.psl.eu}
 
@@
 
-% Let's intereprete this qualitatively.
-% For flat grating
-% the incidence angle $\alpha$
-
 \begin{figure}
@@
 
+\subsection{Python code example}
+
+We provide in the paper repository~\cite{github} a Python code
+to simulate the diffraction effect of a DMD
+by computing the far field pattern for a set of realistic parameters.
+It provides a simple way to estimate the blazed number $\mu$ introduced in eq.~(\ref{eq:blazed_number})
+to assess the quality of the modulation at the desired wavelength
+for a given pixel pitch.
+We also propose an online tool accessible at
+\url{https://www.wavefrontshaping.net/post/id/49}.
+
 % Let's consider first a 1D toy model to have a qualitative understanding of the problem.
@@
 However, in practice, we encounter a substantially distorted pattern,
-like the one represented in Fig.~\ref{fig:dmd_aberr_setup}.d.\\
+like the one represented in Fig.~\ref{fig:dmd_aberr_setup}.d.
+A more detailed depiction of the setup for aberration characterization,
+in the context of a wavefront shaping application in complex media is presented in Fig.~\ref{fig:dmd_aberr_setup2}.a.
+In the case where the medium is a multimode fiber,
+one could leverage the reflection from the input surface
+to directly visualize the intensity pattern of the input plane,
+as shown in Fig.~\ref{fig:dmd_aberr_setup2}.b.\\
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 0.95\textwidth]{images/aberr_setup.pdf}
+  \caption{
+    \textbf{Detailed aberration characterization setup.}
+    (a) A laser beam is expanded and collimated using a telescope (lenses L and L').
+    The DMD is represented in transmission for simplicity.
+    Light reflected from the DMD is focused by a lens L$_\text{1}$
+    onto the complex medium to study, namely a scattering medium (SC)
+    or a multimode fiber (MMF).
+    A beamsplitter is used to image the far field of the DMD onto a camera
+    using a lens (L$_\text{2}$).
+    The intensity pattern, up to a homotetic transformation,
+    represent the input excitation on the mediun
+    which also correspond to the PSF of the illumination setup.
+    (b) In the case of a multimode fiber,
+    one can take advantage of the reflection of the input facet
+    to directly image the intensity pattern of the input plane.
+  }
+  \label{fig:dmd_aberr_setup2}
+\end{figure}
 
@@
 
-In this section, we provide Python code snippets for performing aberration correction.
-We presuppose that phase modulation can be controlled using a function,
-!display_phase_mask(phase_pattern)!, to display a desired phase pattern
-with real values between $0$ and $2\pi$.
-It is important to note that phase modulation utilizing a DMD demands an appropriate setup
-and process. Although not detailed here,
-this is elaborated upon in the referenced tutorial~\cite{Gutierrez2024DMD}.
-Additionally, we assume the existence of a function, !get_camera_image()!,
-which returns the intensity pattern as measured by the camera.
-For generating Zernike polynomials, we make use of the !aotools! package~\cite{Townson2019aotools}.
+We provide in the paper repository~\cite{github} a Python code example
+to simulate the effect of aberration on a DMD
+and then perform a sequential optimization as previously proposed
+to learn the characterize pattern.
+We make use of the !aotools! package~\cite{Townson2019aotools} for generating Zernike polynomials.
+
+% In this section, we provide Python code snippets for performing aberration correction.
+% We presuppose that phase modulation can be controlled using a function,
+% !display_phase_mask(phase_pattern)!, to display a desired phase pattern
+% with real values between $0$ and $2\pi$.
+% It is important to note that phase modulation utilizing a DMD demands an appropriate setup
+% and process. Although not detailed here,
+% this is elaborated upon in the referenced tutorial~\cite{Gutierrez2024DMD}.
+% Additionally, we assume the existence of a function, !get_camera_image()!,
+% which returns the intensity pattern as measured by the camera.
 

### 2023-11-20 python/shrink_data.py

@@ -1,0 +1,17 @@
+import pickle
+
+DATA_FILE = "./data_correction_Zernike_big"
+NEW_DATA_FILE = "./data_correction_Zernike"
+
+with open(DATA_FILE, "rb") as file:
+    data = pickle.load(file)
+
+new_data = {
+    "final_last_image": data["final_last_image"],
+    "zernike_coefficients": data["zernike_coefficients"],
+    "correction_complex_mask": data["correction_complex_mask"],
+    "evol_coeff_imgs": data["evol_coeff_imgs"],
+}
+data = new_data
+
+pickle.dump(data, open(NEW_DATA_FILE, "wb"))

### 2023-11-20 paper/main.tex
test CI
correct refs

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 %
@@
 in particular for high-resolution imaging microscopy
-requiring fast scanning or illumination shaping~\cite{Cha2000,Zhuang2020},
+requiring fast scanning or illumination shaping~\cite{cha2000nontranslational,Zhuang2020},
 biolithography~\cite{yoon2018emerging},
-and optical tweezers~\cite{Gauthier2016}.
+and optical tweezers~\cite{gauthier2016direct}.
 However, DMDs are restricted to hardware binary amplitude modulation and are not optimized for coherent light applications.
@@
 DMDs gained prominence in the 1990s for video projection applications since the 1990s
-under the commercial name of Digital Light Processing (DLP)~\cite{Hornbeck1997,Dudley2003emerging}.
+under the commercial name of Digital Light Processing (DLP)~\cite{hornbeck1997digital,Dudley2003emerging}.
 The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
@@
 
-In the supplementary materials~\cite{}, we present an animation illustrating the dynamic pattern.
+In the supplementary materials~\cite{SI}, we present an animation illustrating the dynamic pattern.
 In off-axis holography,

### 2023-11-21 paper/main.tex
add md

@@
     \textbf{Flat grating vs blazed grating.}
-    ...
+    Far field diffraction patterns for a 1D flat grating (left) and a 1D blazed grating (right)
+    for an input angle of $\alpha = -20^\circ$,
+    a filling fraction of 95\% (corresponding to a 2D filling fraction of $\approx 90$\%),
+    a pixel tilt angle of $\theta_B = 5^\circ$,
+    and a wavelength to pixel pitch ratio $0.05$.
+    Vertical lines represent the angles of the diffraction orders
+    and the black dashed curve represent the amplitude of the field.
   }
@@
 
+\subsection{Dispersion}
+
+DMD are based on small mirror whose response is little dependent on the wavelength,
+which is convenient for broadband applications requiring amplitude modulation.
+However, in
+
 \begin{tldr}

### 2023-11-21 paper/main.tex
add dispersion discussion

@@
 
-\subsection{1D model}
+\subsection{A 1D model}
 
@@
 
-\subsection{2D case}
+\subsection{Tje 2D case}
 
@@
 
-DMD are based on small mirror whose response is little dependent on the wavelength,
-which is convenient for broadband applications requiring amplitude modulation.
-However, in
+DMDs are composed of metallic small mirrors,
+the response of which is minimally affected by wavelength changes.
+This is particularly advantageous for broadband applications requiring amplitude modulation
+and operating on a plane conjugated to the DMD's surface.
+This is the case for the originally intended application of video projection.
+However, for wavefront-shaping applications,
+it is typically required to select a specific diffraction order
+to acheive phase or complex modulation~\cite{lee1979binary,Gutierrez2024DMD}.
+Under such circumstances,
+the wavelength-dependency of the diffraction effect becomes important.
+The blazed number, denoted by $\mu$ (according to Eq.(\ref{eq:blazed_number})),
+scales inversely with the wavelength.
+Fig.~\ref{fig:dispersion} shows the blazed number $\mu$
+as a function of the wavelength for two pixel pitches $d=7.6$\textmu m and $d=10.8$\textmu m,
+with an incident angle of $\alpha' = 20^\circ$.
+Within the visible spectrum, $\mu$ fluctuates between $0$ to $1$ over a typical range of roughly 100 nm.\\
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width = 0.75\textwidth]{images/blazing_number_VS_wavelength.pdf}
+  \caption{
+    \textbf{Dispersion of the diffraction effect.}
+    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident wavelength
+    for an incident angle $\alpha' = 20^\circ$
+    and a pixel pitch of $d=7.6$\textmu m (blue curve)
+    and $d=10.8$\textmu m (dashed black curve).
+  }
+  \label{fig:dispersion}
+\end{figure}
 

### 2023-11-21 paper/main.tex
Updates from Overleaf

--- 2023-11-20 paper/main.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 %
@@
 in particular for high-resolution imaging microscopy
-requiring fast scanning or illumination shaping~\cite{cha2000nontranslational,Zhuang2020},
+requiring fast scanning or illumination shaping~\cite{Cha2000,Zhuang2020},
 biolithography~\cite{yoon2018emerging},
-and optical tweezers~\cite{gauthier2016direct}.
+and optical tweezers~\cite{Gauthier2016}.
 However, DMDs are restricted to hardware binary amplitude modulation and are not optimized for coherent light applications.
@@
 DMDs gained prominence in the 1990s for video projection applications since the 1990s
-under the commercial name of Digital Light Processing (DLP)~\cite{hornbeck1997digital,Dudley2003emerging}.
+under the commercial name of Digital Light Processing (DLP)~\cite{Hornbeck1997,Dudley2003emerging}.
 The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
@@
 In order to utilize Eq.(\ref{eq:blazed_eq}),
-one first needs projecting
+one first needs to project
 the different angles of the problem onto the incident planes
@@
 We can quantify how close we are to the ideal case,
-i.e., when satifying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
+i.e., when satisfying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
 by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:
@@
 and $\mod{2}$ the modulo 2 operation.
-$\mu$ is maximal and equals $1$ when the blazing equation is satisfield,
+$\mu$ is maximal and equals $1$ when the blazing equation is satisfied,
 i.e. when one order of diffraction contains most of the energy,
-and minimal when we are in the worst case scenario,
+and minimal when we are in the worst-case scenario,
 i.e. when four orders of diffraction have a significant and equal intensity.\\
@@
     and $d=10.8$\textmu m (right).
-    Corrseponding far field diffraction pattern (bottom)
+    Corresponding far field diffraction pattern (bottom)
     for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
@@
   \textbf{TL;DR:}\\
-  DMDs act as a blazed gratings.
+  DMDs act as blazed gratings.
   For a given operation wavelength,
@@
 While only capable of providing hardware binary amplitude modulation,
-DMDs serve as an potent tool for wavefront shaping and sensing.
+DMDs serve as a potent tool for wavefront shaping and sensing.
 These applications  critically require characterization and correction of aberrations
@@
 It is noteworthy that when utilized for intensity modulation on a plane conjugated to the DMD plane,
-as it is the done in digital projectors,
+as it is done in digital projectors,
 the system becomes insensitive to the aberrations caused by the DMD surface.
 Consequently, these effects are commonly overlooked and rarely documented in the information provided by the manufacturers.
-They are nonetheless non-negligible, with deviations from a flat surface
-that typically spans across several wavelengths~\cite{Brown2021multicolor}.\\
+Nonetheless, they are non-negligible, with deviations from a flat surface
+that typically spans several wavelengths~\cite{Brown2021multicolor}.\\
 
@@
 the phase pattern of the DMD aberrations in the plane of the modulator.
-Typically, this involves the use of a model for the aberrations,
+Typically, this involves using a model for the aberrations,
 tweaking the parameters to align with the measurements~\cite{Matthes2019Optical,Scholes2019structured, Brown2021multicolor}.
@@
 
-We assume that the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
+We assume the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
 This implies that the modulator can be divided into $N$ sections,
@@
     The intensity pattern, up to a homotetic transformation,
-    represent the input excitation on the mediun
-    which also correspond to the PSF of the illumination setup.
+    represent the input excitation on the medium
+    which also corresponds to the PSF of the illumination setup.
     (b) In the case of a multimode fiber,
@@
 starting from the radial degree $2$.
-Indeed, the initial one, known as the piston, has no influence on the PSF quality.
+Indeed, the initial one, known as the piston, does not influence the PSF quality.
 The subsequent ones, the tip and the tilt, cause the PSF to shift.
@@
   \caption{
-    \textbf{Effect of Zernike polynomials on the optimizaton of the PSF.}
+    \textbf{Effect of Zernike polynomials on the optimization of the PSF.}
     (a) Strehl ratio of the point spread function as a function of the number of Zernike polynomials used
@@
 
-Unlike the original purpose of the DMD, i.e. ampltiude modulation for video projectors,
+Unlike the original purpose of the DMD, i.e. amplitude modulation for video projectors,
 typical scientific applications require a high stability of the generated wavefront.
@@
 where a small change in the phase front can lead to a large change in the output intensity profile.
-While LC-SLMs design has been improved and adapted to scientific applicaitons
+While LC-SLMs design has been improved and adapted to scientific applications
 over the last decades,
-DMD are still a relatively new tools for wavefront shaping and sensing
-and are prone to instabilities that need to be adressed by the user.
-In this section we present the effect of mechanical and thermal instabilities
+DMD are still relatively new tools for wavefront shaping and sensing
+and are prone to instabilities that need to be addressed by the user.
+In this section, we present the effect of mechanical and thermal instabilities
 and how to limit their impact on the wavefront quality with simple solutions.\\
@@
 
-In the supplementary materials~\cite{SI}, we present an animation illustrating the dynamic pattern.
+In the supplementary materials~\cite{}, we present an animation illustrating the dynamic pattern.
 In off-axis holography,
@@
 with a displacement equivalent to the period of the fringes corresponding to $2\pi$~\cite{Cuche2000spatial}.
-This permits us to estimate the fluctuation in phase over time at a given postition of the output plane.
+This permits us to estimate the fluctuation in phase over time at a given position of the output plane.
 As illustrated in Fig.~\ref{fig:phase_vibrations} (depicted by the red curve),
@@
 
-An simple yet effective solution consists in dampening the vibrations
+A simple yet effective solution consists in dampening the vibrations
 at the flat cable's level by clamping it with a soft material,
@@
 Specifically, the chips heats up more quickly when increasing the frame rate.
-The increase of temperature can reach more that $15$ degrees Celsius
+The increase of temperature can reach more than $15$ degrees Celsius
 when running a sequence at maximum speed (20-30 kHz)~\cite{Rudolf2021thermal}.
@@
 Temperature fluctuations can cause deformations on the chip's surface
-and can modificy the phase response of the glass protective window.
+and can modify the phase response of the glass protective window.
 This creates low order aberrations, which degrade the quality of the wavefront.
@@
 to directly measure the effect on the studied system's response.
-To do so, we use a setup to the one presented in the previous section
+To do so, we use a setup similar to the one presented in the previous section
 and depicted in Fig.~\ref{fig:MMF_ref}.
 We can then estimate the field or intensity decorrelation over time.
-We present here results with intesity correlation, as it does not require a reference arm.
+We present here results with intensity correlation, as it does not require a reference arm.
 The measured output pattern typically take the form of a seemingly random speckle pattern,
-that is sensitive to minute changes of the input wavefront.
+that is sensitive to minute changes in the input wavefront.
 The correlation estimation is obtained by comparing the output intensity pattern
@@
 The more efficient way to counteract this effect is to use a
-close-loop system to stabilize the temperature of the DMD chip.
+closed-loop system to stabilize the temperature of the DMD chip.
 This can be achieved by using a thermoelectric cooler
@@
   DMDs take about an hour to thermaly stabilize when a sequence is running.
-  It can conuntered by using a thermoelectric cooler
+  It can countered by using a thermoelectric cooler
   and a temperature sensor.
-  It can also be simply mitigated by letting a sequence running for few hours
+  It can also be simply mitigated by letting a sequence run for few hours
   before starting the experiment.

### 2023-11-21 paper/main.tex

@@
 
+\subsection{Dispersion}
+
+DMDs are composed of metallic small mirrors,
+the response of which is minimally affected by wavelength changes.
+This is particularly advantageous for broadband applications requiring amplitude modulation
+and operating on a plane conjugated to the DMD's surface.
+This is the case for the originally intended application of video projection.
+However, for wavefront-shaping applications,
+it is typically required to select a specific diffraction order
+to acheive phase or complex modulation~\cite{lee1979binary,Gutierrez2024DMD}.
+Under such circumstances,
+the wavelength-dependency of the diffraction effect becomes important.
+The blazed number, denoted by $\mu$ (according to Eq.(\ref{eq:blazed_number})),
+scales inversely with the wavelength.
+Fig.~\ref{fig:dispersion} shows the blazed number $\mu$
+as a function of the wavelength for two pixel pitches $d=7.6$\textmu m and $d=10.8$\textmu m,
+with an incident angle of $\alpha' = 20^\circ$.
+Within the visible spectrum, $\mu$ fluctuates between $0$ to $1$ over a typical range of roughly 100 nm.\\
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width = 0.75\textwidth]{images/blazing_number_VS_wavelength.pdf}
+  \caption{
+    \textbf{Dispersion of the diffraction effect.}
+    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident wavelength
+    for an incident angle $\alpha' = 20^\circ$
+    and a pixel pitch of $d=7.6$\textmu m (blue curve)
+    and $d=10.8$\textmu m (dashed black curve).
+  }
+  \label{fig:dispersion}
+\end{figure}
+
 \begin{tldr}

### 2023-11-28 paper/main.tex
add Rodrigo's comment + conclusion

@@
 under the commercial name of Digital Light Processing (DLP)~\cite{hornbeck1997digital,Dudley2003emerging}.
-The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.\\
+The technology enables high-resolution, high-speed, and high-contrast-ratio modulation of light.
 DMDs operate by toggling the state of small mirrors between two distinct angles, denoted as $\pm \theta_\text{DMD}$.
@@
 where $\lambda$ is the wavelength of the light,
-and $\theta_D$ is the angle of the first-order diffraction.
-The intensity of the individual diffraction orders is influenced by both the size and the response of a single pixel.
+$\theta_D$ is the angle of the first-order diffraction,
+and $p$ is an integer value denoting the orders of diffraction.
+The intensity of the individual diffraction orders is influenced by
+the response of a single pixel,
+constituting the unit cell of the grating,
+and that is governed by $d$, $d'$, and its rotation angle in the case of a DMD.
 Importantly, we can decouple the effects of the grating's periodicity,
@@
 For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
-we can assume that the response of a single pixel is uniform its surface.
+we can assume that the response of a single pixel is uniform over its surface.
 Similarly, in the scenario of a blazed grating,
@@
 relative to the case of normal incidence,
-while the blazed angle $\theta_B$—the tilt angle of the mirror in the DMD—
+while the blazed angle $\theta_B$
+—the tilt angle of the mirror in the DMD projected onto the axis we consider—
 only shifts the envelope by an angle of $2\theta_B$.
@@
 this corresponds to the angle of the zeroth-order diffraction,
-which nears zero for all other orders.
+and the intensity nears zero for all other orders.
 In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
@@
 this position may not align with a diffraction order anymore~\cite{Park2015properties}.
-
-\begin{optional}
-  Assuming the effect of the device's finite size and illumination to be negligible,
-  and situating ourselves within the context of the small-angle approximation,
-  we can represent the field reflected from the device
-  under the influence of plane wave illumination
-  across two systems as follows:
-
-  \begin{equation}
-    \begin{aligned}
-      R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
-      R_\text{blazed}(x) & \propto
-      \underbrace{
-        \left(
-        \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
-        \times
-      \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle}                                         \\ \text{+ angle of incidence}}}
-        %\text{blazed angle + angle of incidence}
-        \right)
-      }_\text{pixel response}
-      \otimes_x
-      \left[
-        \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
-        .
-        \underbrace{
-          e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
-        }_\text{angle of incidence}
-        \right]
-      \, ,
-    \end{aligned}
-    \label{eq:grating_response}
-  \end{equation}
-
-  with $\Pi(x)$ the rectangular function,
-  representing the finite size of the pixel,
-  defined as:
-
-  \begin{equation}
-    \Pi(x) =
-    \begin{cases}
-      1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
-      0, & \text{otherwise}.
-    \end{cases}
-  \end{equation}
-
-  The intensity as a function of the angle in the far-field
-  is given, up to a homotetic transformation,
-  by the absolute value squared of
-
-  The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
-
-  \begin{equation}
-    \begin{aligned}
-      I_\text{flat}(\theta) \propto
-      \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
-      \times
-      \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
-      I_\text{blazed}(\theta) \propto
-      \underbrace{
-        \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D)
-      }_\text{orders of diffraction}
-      \times
-      \underbrace{
-        \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
-      }_\text{envelope} \, .
-    \end{aligned}
-  \end{equation}
-
-\end{optional}
-A more accurate computation of the far field can be found  in ~\cite{Wang2023diffraction}.
-We observe that the envelope (right hand term) is maximal for $\text{sin}(\theta_{max}) = \text{sin}(\alpha-2\theta_B)$
-while the effect of the periodicity (left hand term) is maximal for $\text{sin}(\theta_p)+ \text{sin}(\alpha) = p \lambda/d$.
-with $p$ an integer value.
+We provide in Appendix A.  a 1D computation of this effect.
+A more accurate computation of the far field can be found  in ~\cite{Wang2023diffraction}.\\
+
 We represent in Fig.~\ref{fig:gratings} the angular response of a flat grating and a blazed grating
 for a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
-For a flat grating, only the zero-th order has a significant and comparable intensity.
-For a blazed grating, in this example close to the worst case scenario,
-two diffraction orders have a significant intensity,
+For a flat grating, the zero-th order contains most of the intensity,
+the other orders being negligible in comparison.
+For the blazed grating example shown,
+we are in a situation close to the worst case scenario:
+Two diffraction orders have a significant and comparable intensity,
 and other orders also have a non-negligible contributions.
@@
   \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
-  = 2 \,\text{sin}(\theta_B)  \,\text{sin}(\theta_B-\alpha)
+  = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
   = p\frac{\lambda}{d} \, .
@@
 \end{equation}
+
+\noindent We note that the incident angle $\alpha$ olso affects the diffraction efficiency.
 
@@
 
-\subsection{Tje 2D case}
+\subsection{The 2D case}
 
@@
 it is preferable to work with the incident and outgoing beams
-which have the optical axis contained in the horizontal plane.
+which have the optical axis contained in the horizontal plane,
+i.e. a plane parallel to the table surface.
 A straightforward and prevalent solution is to rotate the chip by 45 degrees
@@
 which aligns the pixel’s axis of rotation to be vertical.
-This configuration is depicted in Fig.~\ref{fig:2d_geom}b.
+This configuration is depicted in Fig.~\ref{fig:2d_geom}b.\\
 
@@
     (a) Schematic representation of the geometry of the DMD.
+    The incident  and reflection angles $\alpha'$ and $\theta_0'$ are situated
+    within the horizontally plane, illustrated in yellow.
     (b) Photograph of the DMD chip oriented so that the rotation axis of the pixels
@@
 $\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$,
-$\beta = \text{arctan}\left(\text{tan}(\beta')/\sqrt{2}\right)$,
+$\theta = \text{arctan}\left(\text{tan}(\theta')/\sqrt{2}\right)$,
 and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$,
 where $\theta_\text{DMD}$ represents the true rotation angle of the mirrors relative to the diagonal of the pixels,
-and $\alpha'$ and $\beta'$ denote the incident and outgoing angles in the horizontal plane.
+and $\alpha'$ and $\theta'$ denote the incident and outgoing angles in the horizontal plane.
 We can quantify how close we are to the ideal case,
@@
   \left|
-  \lfloor 2 \frac{d}{\lambda}
+  \lfloor 4 \frac{d}{\lambda}
   \left[
-    \text{sin}(\alpha)  +\text{sin}(\beta)
+    \text{sin}(\theta_B)\text{cos}(\theta_B-\alpha)
     \right]
@@
 
-with $\lfloor . \rfloor$ representing the integer part,
+with $\lfloor . \rfloor$ representing the floor function,
 and $\mod{2}$ the modulo 2 operation.
@@
 Fig.~\ref{fig:mu76} shows the estimated blazed number $\mu$
-as a function of the angle of incidence $\alpha'$,
+as a function of the angle of incidence $\alpha'$ in the horizontal plane,
 along with the far field diffraction pattern for two distinct incident angles.
 It should be noted that the efficiency of diffraction can be altered by adjusting
-the angle of incidence, yet its impact is relatively confined
+the angle of incidence.
+However, its impact is relatively confined
 within an acceptable angular range that aligns with experimental limitations
@@
 and therefore, will contribute as an interference to the modulated wavefront.
+We show in Fig.~\ref{fig:xtalk} the normalized amplitude of the diffraction patterns
+corresponding to all the pixels in the {\em on} (blue curve) and {\em off} (orange curve) states
+for the same experimental conditions but with two different pixel pitches
+leading to situations close to the worst (a) and best (b) case scenarios.
+We observe the presence of a non-negligible contribution of the {\em off} state
+at the main diffraction order of the {\em on} state in the first case.
 While this contribution might appear weak,
@@
 
-the best correction if obtained for
+\noindent the best correction is obtained for
 
@@
 
-% \section{Control Vialux DMDs with Python}
-
-% \cite{hofling2004alp}
-
-% \cite{popoff2016alp4lib}
-
-\section{Acknowledgments}
-Authors wishing to acknowledge assistance or encouragement from
-colleagues, special work by technical staff or financial support from
-organizations should do so in an unnumbered `Acknowledgments' section
-immediately following the last numbered section of the paper. In \verb"iopart.cls" the
-command \verb"\ack" sets the acknowledgments heading as an unnumbered
-section.
+\section{Conclusion}
+DMDs are powerful tools for wavefront shaping applications in complex media
+due to their high pixel count, relative low cost, and high refresh rate.
+However, mostly due to their original purpose of amplitude modulation
+of incoherent light for video projection,
+several effects need to be taken into account when using DMDs for wavefront shaping.
+First, the choice of its pixel pitch must be made carefully
+by considering the wavelength of operation
+to ensure a good modulation quality and diffraction efficiency.
+Furhtermore, the DMD surface is not flat and can introduce aberrations
+which can be corrected using a simple optimization procedure.
+Finally, the DMD is sensitive to mechanical vibrations and thermal variations,
+which can be mitigated ensuring a good mechanical and thermal isolation of the device.
+
+% \section{Acknowledgments}
+
+\section*{Appendix A: 1D calculation of the DMD diffraction effect}
+\label{app:1D}
+
+Assuming the effect of the device's finite size and illumination to be negligible,
+and situating ourselves within the context of the small-angle approximation,
+we can represent the field reflected from the device
+under the influence of plane wave illumination
+across two systems as follows:
+
+\begin{equation}
+  \begin{aligned}
+    R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
+    R_\text{blazed}(x) & \propto
+    \underbrace{
+      \left(
+      \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
+      \times
+    \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle}                                         \\ \text{+ angle of incidence}}}
+      \right)
+    }_\text{pixel response}
+    \otimes_x
+    \left[
+      \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
+      .
+      \underbrace{
+        e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
+      }_\text{angle of incidence}
+      \right]
+    \, ,
+  \end{aligned}
+  \label{eq:grating_response}
+\end{equation}
+
+with $\Pi(x)$ the rectangular function,
+representing the finite size of the pixel,
+defined as:
+
+\begin{equation}
+  \Pi(x) =
+  \begin{cases}
+    1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
+    0, & \text{otherwise}.
+  \end{cases}
+\end{equation}
+
+The intensity as a function of the angle in the far-field
+is given, up to a homotetic transformation,
+by the absolute value squared of
+
+The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
+
+\begin{equation}
+  \begin{aligned}
+    I_\text{flat}(\theta) \propto
+    \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
+    \times
+    \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
+    I_\text{blazed}(\theta) \propto
+    \underbrace{
+      \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D)
+    }_\text{orders of diffraction}
+    \times
+    \underbrace{
+      \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
+    }_\text{envelope} \, .
+  \end{aligned}
+\end{equation}
+
+We observe that the envelope (right hand term) is maximal for $\text{sin}(\theta_{max}) = \text{sin}(\alpha-2\theta_B)$
+while the effect of the periodicity (left hand term) is maximal for $\text{sin}(\theta_p)+ \text{sin}(\alpha) = p \lambda/d$,
+representing the orders of diffraction.
 

### 2023-11-29 paper/main.tex
add data section
add abstract

@@
 \begin{abstract}
-  Abstract
+  Digital micromirror devices have gained popularity in wavefront shaping,
+  offering a high frame rate alternative to liquid crystal spatial light modulators.
+  They are relatively inexpensive, offer high resolution,
+  are easy to operate, and a single device can be used in a broad optical bandwidth.
+  However, some technical drawbacks must be considered
+  to achieve optimal performance.
+  These issues, often undocumented by manufacturers,
+  mostly stem from the device's original design for video projection applications.
+  Herein, we present a guide to characterize and mitigate these effects.
+  Our focus is on providing simple and practical solutions
+  that can be easily incorporated into a typical wavefront shaping setup.
 \end{abstract}
@@
 
+\section*{Data and code availability}
+
+\noindent Data and code examples are available in the dedicated repository~\cite{github}.
+
 \section*{Appendix A: 1D calculation of the DMD diffraction effect}

### 2024-05-29 paper/main.tex
Updates from Overleaf

@@
 \usepackage{listings}
+\usepackage{xcolor}
 
@@
 \definecolor{light_orange}{RGB}{255,193,134}
+\newcommand{\red}[1]{#1}
 
@@
 Nonetheless, LC-SLMs are hampered by their slow response time,
-permitting only a modulation speed ranging from a few Hz to about 100 Hz.\\
+permitting only a modulation speed ranging from a few Hz to \red{a few hundred} Hz.\\
 
@@
 this tutorial aims to provide a guide for easily setting up a DMD for wavefront shaping applications in complex media.
-In particular, we provide characterization and validation procedures that requires minimal changes compared to typical wavefront shaping setups.\\
+In particular, we provide characterization and validation procedures that require minimal changes compared to typical wavefront shaping setups. 
+\red{
+More specifically, we place ourselves in a standard experiment where the goal is to shape the complex wavefront impinging on a complex medium and control or measure its output response. 
+This is typically the case for a focusing experiment~\cite{Vellekoop2007focusing} or for measuring the transmission matrix of a complex medium~\cite{Popoff2010Measuring}.
+}\\
 We first introduce the diffraction properties of a DMD
@@
 along with the far field diffraction pattern for two distinct incident angles. 
+Fig.~\ref{fig:mu_2d} shows the blazed number as a function of both the incident angle 
+and the wavelength for pixel pitches of $d=7.6$\textmu m and $d=10.8$\textmu m.
 It should be noted that the efficiency of diffraction can be altered by adjusting
@@
 (i.e. for angles far from $\pm 90^\circ$).
-Far field patterns are centered around the maximum of the envelope
+Far-field patterns are centered around the maximum of the envelope
 $\theta_\text{max} = 2\theta_B - \alpha$
@@
 \begin{figure}
-  \centering
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{images/mu_76.pdf}
-  \end{subfigure}
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{images/mu_108.pdf}
-  \end{subfigure}
-  \caption{
-    \textbf{Blazed number and far field diffraction patterns, extreme cases.}
+  % \centering
+  % \begin{subfigure}{0.49\textwidth}
+  %   \centering
+  %   \includegraphics[width = \textwidth]{images/mu_76.pdf}
+  % \end{subfigure}
+  % \begin{subfigure}{0.49\textwidth}
+  %   \centering
+  %   \includegraphics[width = \textwidth]{images/mu_108.pdf}
+  % \end{subfigure}
+  \includegraphics[width = \textwidth]{images/mu_1d.pdf}
+  \caption{
+    \textbf{Blazed number and far-field diffraction patterns, extreme cases.}
     Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha'$  (top)
-    for a pixel pitch of $d=7.6$\textmu m (left)
-    and $d=10.8$\textmu m (right).
-    Corresponding far field diffraction pattern (bottom)
+    for a pixel pitch of $d=7.6$\textmu m (a.)
+    and $d=10.8$\textmu m (b.).
+    Corresponding far-field diffraction pattern (bottom)
     for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
@@
   \label{fig:mu76}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{images/mu_2d.pdf}
+  \caption{
+    \textbf{Blazed number as a function of the wavelength and the incident angle.
+    }
+    Color map representation of the blazed number for a pixel pitch of $d=7.6$\textmu m (a.)
+    and $d=10.8$\textmu m (b.).
+  }
+  \label{fig:mu_2d}
 \end{figure}
@@
 to directly visualize the intensity pattern of the input plane,
-as shown in Fig.~\ref{fig:dmd_aberr_setup2}.b.\\
+as shown in Fig.~\ref{fig:dmd_aberr_setup2}.b. 
+\red{
+Note that for the later approach to be reliable, it is necessary to properly align the system to ensure that the plane of the input facet corresponds exactly to the Fourier plane of the DMD, and that the input facet plane is conjugated with the camera plane.
+}
+\\
 
@@
     \textbf{Correction phase map.}
-    (a) Schematic representation of the function we want to optimize.
+    (a) Schematic representation of the function we want to optimize, 
+    i.e. we optimize the ratio between 
+    the sum of the intensity of the measured image in a diffraction-limited disk
+    and the sum of the intensity elsewhere.    
     We both want to maximize the intensity of the main lobe of the point spread function
@@
 
+\red{Empirically, we observe deformations of the pattern 
+at the output of the medium under study when the DMD chip heats up after switching on the device. 
+These deformations originate from a temperature-dependent perturbation of the modulated field. 
+We attribute this effect to the transparent part of the DMD 
+(i.e., the protection glass window) 
+having a refractive index that significantly changes with temperature. 
+Once the device reaches thermal equilibrium, this effect stabilizes, 
+allowing us to find the appropriate correction mask as described in Section 3 of the manuscript. 
+It is important to stress that it is the thermal fluctuations that impact the stability of the modulation, 
+not the temperature itself, 
+as we obtain stable experimental conditions even when equilibrium is reached 
+at the maximal operating temperature of the modulator without thermal cooling.\\}
+
 Before initiating a wavefront shaping experiment,
@@
 to directly measure the effect on the studied system's response. 
+\red{Directly using the output of the complex medium under study presents multiple advantages.
+First, it allows the setup to remain unchanged and facilitates easy re-characterization of the effect if some parameters change.
+Additionally, complex media tend to be very sensitive to input phase changes, with the sensitivity varying from one medium to another.
+By using the complex medium itself, we obtain the exact sensitivity needed for the current setup.
+Therefore, regardless of the qualitative changes in the input field, stability in the output indicates that the setup is adequate for studying the medium in question.}
 To do so, we use a setup similar to the one presented in the previous section
@@
   \textbf{TL;DR:}\\
-  DMDs take about an hour to thermaly stabilize when a sequence is running.
+  DMDs take about an hour to thermally stabilize when a sequence is running.
   It can countered by using a thermoelectric cooler

### 2024-08-01 paper/main.tex

@@
 \definecolor{light_orange}{RGB}{255,193,134}
+% \newcommand{\red}[1]{\textcolor{red}{#1}}
 \newcommand{\red}[1]{#1}
@@
 
-\author{Sébastien M. Popoff~\orcidlink{0000-0002-7199-9814}}
+\author{Sébastien M. Popoff~\footnote{corresponding author}~\orcidlink{0000-0002-7199-9814}}
 \address{Institut Langevin, ESPCI Paris, PSL University, CNRS, France}
@@
     \centering
-    \includegraphics[width = \textwidth]{images/pixel_2.jpg}
+    \includegraphics[width = \textwidth]{images/pixel_2.png}
     \label{fig:pix_right}
@@
   \label{fig:mu_2d}
+\end{figure}
+
+To illustrate the link between the blaze number and the efficiency of the diffraction setup, 
+we compute the ratio of the energy in the brightest diffraction spot to the total diffracted energy 
+for different values of the wavelength, 
+with fixed values for the pixel pitch (7.6 \textmu m) and incident angle (normal incidence).
+We simulate the DMD with no gap between the pixels (i.e., a filling fraction of 100\%) 
+and with 30 pixels in each direction, 
+where all the mirrors of the pixels are in the same state. 
+The results are shown in Fig.~\ref{fig:mu_vs_diff}.
+We observe that the maxima (resp. minima) of the diffraction efficiency 
+correspond to the maxima (resp. minima) of the blaze number. 
+Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \textwidth]{images/mu_vs_diffeff.pdf}
+  \caption{
+    \textbf{Blazed number vs diffraction efficiency.
+    }
+    Plot of the blazed number and the diffraction efficiency for a normal incidence and a pixel pitch of  7.6 \textmu m 
+    as a function of the wavelength. 
+    The diffraction efficiency is estimated by comparing the energy in a diffracted limited spot at the 
+    location of the brightest diffraction order to the total diffracted energy.
+  }
+  \label{fig:mu_vs_diff}
 \end{figure}
@@
 
+\section*{Acknowledgments}
+
+\noindent 
+S.M.P, M.W.M, and R.G.C and acknowledge the French \textit{Agence Nationale pour la Recherche} grants No. ANR-23-CE42-0010-01 MUFFIN and No. ANR-20-CE24-0016 MUPHTA 
+and the  Labex WIFI grant No. ANR-10-LABX-24, ANR-10-IDEX-0001-02 PSL*.
+
 \section{Bibliography}

### 2025-09-30 paper/main.tex

@@
 
+\author{Louis Malosse~\orcidlink{0009-0004-9006-9374}}
+\address{LIRA, Observatoire de Paris, Université PSL, CNRS, Université Paris Cité, Sorbonne Université, France}
+\address{ISMO, University Paris Saclay, CNRS, France}
+\ead{louis.malosse@universite-paris-saclay.fr}
+
 \author{Rodrigo Gutiérrez-Cuevas~\orcidlink{0000-0002-3451-6684}}
@@
 \address{Racah Institute of Physics, The Hebrew University of Jerusalem, Israel}
+
+\author{Jean Commère~\orcidlink{0009-0000-4710-4972}}
+\address{LIRA, Observatoire de Paris, Université PSL, CNRS, Université Paris Cité, Sorbonne Université, France}
+\ead{jean.commere@obspm.fr}
+
+\author{Marie Glanc}
+\address{LIRA, Observatoire de Paris, Université PSL, CNRS, Université Paris Cité, Sorbonne Université, France}
+\ead{marie.glanc@obspm.fr}
+
+\author{Raphaël Galicher}
+\address{LIRA, Observatoire de Paris, Université PSL, CNRS, Université Paris Cité, Sorbonne Université, France}
+\ead{raphael.galicher@obspm.fr}
 
@@
 A significant distinction between liquid crystal modulators and DMDs
-lies in the geometry of the pixel surface.
+lies in the geometry of the pixel surface: 
+The pixels of the DMD are not flat, but are mirrors that can be tilted 
+in two different positions around a rotation axis
+along the diagonal of the square shape of the pixels.
 This difference gives rise to diffraction effects that can adversely affect
-both the modulation quality and system efficiency.
-The impact of these diffraction effects is highly contingent on several factors:
-the wavelength of the illumination, the pixel pitch, and both the incident and outgoing angles.
-Therefore, alongside selecting an appropriate anti-reflection coating,
+both modulation quality and system efficiency.
+The impact of these diffraction effects is highly dependent on several factors:
+the wavelength of illumination, the pixel pitch, and both the incident and outgoing angles.
+Therefore, in conjunction with selecting an appropriate anti-reflection coating,
 it is crucial to ensure that the pixel pitch is compatible with the specific configuration being used.
@@
 we consider a 1D array of pixels as illustrated in Fig.~\ref{fig:grating_geom}.
-% For simplicity, we will focus on scenarios where both the incident and outgoing angles 
-% are nearly perpendicular to the modulator surface, and the mirror tilt angle is minimal.
 Initially, let's assume that all pixels are in the same state and are illuminated by a plane wave originating from the far field.
@@
 In general, a grating gives rise to various diffraction orders
-with differing intensities
+with different intensities
 and angles $\theta_p$, as dictated by the grating equation:
 $
-  \text{sin}(\theta_p) = p\lambda/d = p \, \text{sin}(\theta_D),
+ \text{sin}(\theta_p)+ \text{sin}(\alpha) = p\lambda/d = p \, \text{sin}(\theta_D),
 $
-where $\lambda$ is the wavelength of the light,
-$\theta_D$ is the angle of the first-order diffraction,
+where $\lambda$ is the wavelength of the light, $\alpha$ is the incident angle, 
+$\theta_D = \arcsin \left(\lambda/S\right)$ is the angle of the first diffraction order,
 and $p$ is an integer value denoting the orders of diffraction.
@@
 This results in a trivial shift of the angular diffraction pattern by an angle $\alpha$.
-In essence, the incident angle $\alpha$ serves to shift the entire angular diffraction pattern
+The incident angle $\alpha$ has the effect of shifting the entire angular diffraction pattern
 relative to the case of normal incidence,
 while the blazed angle $\theta_B$
-—the tilt angle of the mirror in the DMD projected onto the axis we consider—
+—the tilt angle of the micro-mirrors of the DMD—
 only shifts the envelope by an angle of $2\theta_B$.
@@
 this corresponds to the angle of the zeroth-order diffraction,
-and the intensity nears zero for all other orders.
+and the intensity is comparatively small for all other orders.
 In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
 The addition of a blazed angle $\theta_B$ results in both
-a shift in the envelope and in the position of its maximum,
-now indicated by $\theta_0 = 2\theta_B - \alpha$.
+a shift in the envelope and in the position of its maximum 
+$\theta_0 = 2\theta_B - \alpha$.
 In the general case,
@@
 Two diffraction orders have a significant and comparable intensity,
-and other orders also have a non-negligible contributions.
+and other orders also have non-negligible contributions.
 In the optimal scenario, where the peak of the envelope corresponds to a diffraction order,
@@
 
-\noindent We note that the incident angle $\alpha$ olso affects the diffraction efficiency.
-
 \begin{figure}
@@
     a pixel tilt angle of $\theta_B = 5^\circ$,
-    and a wavelength to pixel pitch ratio $0.05$.
-    Vertical lines represent the angles of the diffraction orders
-    and the black dashed curve represent the amplitude of the field.
+    and a wavelength to pixel pitch ratio $\lambda/d=0.05$.
+    Vertical lines represent the angles of the diffraction orders, peaked at $\theta_p$ which satisfy $\sin(\theta_p)+\sin(\alpha)=p\lambda/d$, 
+    and the black dashed curve represents the amplitude of the field which peaks at $\theta_0=2\theta_B-\alpha$.
   }
@@
 \end{figure}
-
-% \begin{equation}
-%   R_\text{flat}(x) \propto \Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d) 
-% \end{equation}
-
-% small angle approximation
-% \begin{equation}
-%   R_\text{blazed}(x) \propto \left[\Pi\left(x/d'\right) \times e^{j\frac{2\pi}{\lambda}\theta_B}\right]\otimes_x \sum_k \delta(x-k d) 
-% \end{equation}
 
@@
 (refer to Fig.~\ref{fig:2d_geom}a).
-Pixels are uniformly repeated along these axes.
+The pixels are uniformly repeated along these axes.
 However, a technical challenge arises
@@
 i.e. a plane parallel to the table surface.
-A straightforward and prevalent solution is to rotate the chip by 45 degrees
+A straightforward and common solution is to rotate the chip by 45 degrees
 relative to the horizontal plane,
-which aligns the pixel’s axis of rotation to be vertical.
+which aligns the pixel axis of rotation to be vertical.
 This configuration is depicted in Fig.~\ref{fig:2d_geom}b.\\
@@
     (a) Schematic representation of the geometry of the DMD.
-    The incident  and reflection angles $\alpha'$ and $\theta_0'$ are situated
+    The incident  angle $\alpha$ and the reflection angle $\theta_0$ (defined by the peak of the diffraction envelope) are situated
     within the horizontally plane, illustrated in yellow.
@@
 
-The 2D array can be seen as 2 orthogonal 1D gratings in the $x$ and $y$ directions,
-both having the same properties.
-A more comprehensive depiction of the 2D system can be found in~\cite{Scholes2019structured}.
-In order to utilize Eq.(\ref{eq:blazed_eq}),
-one first needs to project
-the different angles of the problem onto the incident planes
-of the two 1D gratings.
-This yields
-$\alpha = \text{arctan}\left(\text{tan}(\alpha')/\sqrt{2}\right)$,
-$\theta = \text{arctan}\left(\text{tan}(\theta')/\sqrt{2}\right)$,
-and $\theta_B = \text{arctan}\left(\text{tan}(\theta_\text{DMD})/\sqrt{2}\right)$,
-where $\theta_\text{DMD}$ represents the true rotation angle of the mirrors relative to the diagonal of the pixels,
-and $\alpha'$ and $\theta'$ denote the incident and outgoing angles in the horizontal plane.
+A more detailed description of the 2D system 
+and computation of its response 
+can be found in~\cite{Scholes2019structured}.
+However, we can find a simple condition for
+having most of the energy in one diffraction order 
+by looking at the effect of the tilts of the mirrors 
+when the axis is rotated by 45 degrees compared to 
+the axis of the pixel array (axes $x$ and $y$). 
+We consider a plane wave with an incoming angle $\alpha$ contained in the horizontal plane 
+(i.e. orthogonal to the rotation axis of the micro-mirrors).
+The effect of this angle is represented by a phase term of the form:
+
+\begin{equation}
+{\vec{k}_{in}\cdot\vec{r}} = {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}}
+\, ,
+\end{equation}
+
+with $\vec{k}_{in}$ the incoming wavevector and $\vec{r}$ the position in the modulator plane, 
+represented by the coordinate $(x,y)$.
+
+The grating condition is achieved when the output contributions are in phase 
+for the output angle corresponding to the specular reflection on the tilted mirrors. 
+It corresponds to the plot on the left in Fig.\ref{fig:gratings} for the 1D case.
+The phase contribution resulting from an output angle $2\theta_B-\alpha$, 
+corresponding to the specular reflection on each micro-mirror,
+reads:
+
+\begin{equation}
+\vec{k}_{out}\cdot\vec{r} 
+= {\frac{2\pi}{\lambda}\sin\left(2\theta_B - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}}
+\, ,
+\end{equation}
+
+The total phase then reads:
+
+\begin{equation}
+\phi(x,y) = \vec{k}_{in}.\vec{r}  + \vec{k}_{out}.\vec{r} =  
+{\frac{2\pi}{\lambda}\sin\left(2\theta - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}} +
+{\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}} 
+\, .
+\end{equation}
+
+We see that $x$ and $y$ have the same effect.
+The grating condition is then achieved when two consecutive pixels, 
+separated by a distance $d$ in both the $x$ and $y$ directions, are in phase.
+I.e. for
+
+\begin{equation}
+  \phi(x,y) = \phi(x+d,y) = \phi(x,y+d) \, .
+\end{equation}
+
+The new blazed grating equation then reads~:
+
+\begin{equation}
+  \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
+  = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
+  = p\frac{\sqrt{2}\lambda}{d} \, ,
+  \label{eq:blazed_eq_new}
+\end{equation}
+
+with $p$ an integer.
+Note that this is what one would obtain using Eq.~\ref{eq:blazed_eq} for 
+blazed grating with a pitch $d/\sqrt{2}$, or a 1D grating rotated by 45 degrees.
+Those systems are not equivalent in the general case, 
+but the grating condition in the 2D case is
+similar when considering an incident plane wave in the horizontal plane.\\
+
 We can quantify how close we are to the ideal case,
-i.e., when satisfying the blazed equation outlined by Eq.(\ref{eq:blazed_eq}),
+i.e., when satisfying the blazed this equation,
 by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:
@@
   \mu =
-  \left|
-  \lfloor 4 \frac{d}{\lambda}
+  \left| 4 \frac{d}{\sqrt{2}\lambda}
   \left[
@@
     \right]
-  \rfloor
   \mod{2} -1
@@
 
-with $\lfloor . \rfloor$ representing the floor function,
-and $\mod{2}$ the modulo 2 operation.
+where $\mod{2}$ represents the modulo 2 operation.
 $\mu$ is maximal and equals $1$ when the blazing equation is satisfied,
@@
 
-To demonstrate the effect, we conduct a simulation of a DMD using Python
+To demonstrate the effect, we first conduct a simulation of a DMD using Python
 (refer to tutorial and code in~\cite{WFSnet_diffraction})
@@
 Fig.~\ref{fig:mu76} shows the estimated blazed number $\mu$
-as a function of the angle of incidence $\alpha'$ in the horizontal plane,
+as a function of the angle of incidence $\alpha$ in the horizontal plane,
 along with the far field diffraction pattern for two distinct incident angles. 
@@
 It should be noted that the efficiency of diffraction can be altered by adjusting
-the angle of incidence.
+the angle of incidence $\alpha$.
 However, its impact is relatively confined
@@
 Far-field patterns are centered around the maximum of the envelope
-$\theta_\text{max} = 2\theta_B - \alpha$
+for an output angle $2\theta_B - \alpha$
 (marked by a yellow cross).
-We see that for small values of $\alpha$,
+We see that for small positive values of $\alpha$,
 the pixel pitch of $d=10.8$\textmu m leads to a blazed number $\mu$
-close $1$.
+close to $1$.
 It corresponds in the far field to having one bright order of diffraction
@@
 \begin{figure}
-  % \centering
-  % \begin{subfigure}{0.49\textwidth}
-  %   \centering
-  %   \includegraphics[width = \textwidth]{images/mu_76.pdf}
-  % \end{subfigure}
-  % \begin{subfigure}{0.49\textwidth}
-  %   \centering
-  %   \includegraphics[width = \textwidth]{images/mu_108.pdf}
-  % \end{subfigure}
   \includegraphics[width = \textwidth]{images/mu_1d.pdf}
   \caption{
-    \textbf{Blazed number and far-field diffraction patterns, extreme cases.}
-    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha'$  (top)
+    \textbf{Blazed number and far-field diffraction patterns.}
+    Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident angle $\alpha$  (top)
     for a pixel pitch of $d=7.6$\textmu m (a.)
     and $d=10.8$\textmu m (b.).
-    Corresponding far-field diffraction pattern (bottom)
-    for two incident angles $\alpha' = 0^\circ$ and $\alpha' = 25^\circ$.
-    The yellow cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
+    Corresponding far-field diffraction patterns (bottom)
+    for two incident angles $\alpha = -15^\circ$ and $\alpha = 0^\circ$.
+    The red cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
   }
@@
 We simulate the DMD with no gap between the pixels (i.e., a filling fraction of 100\%) 
-and with 30 pixels in each direction, 
+and with 10 pixels in each direction, 
 where all the mirrors of the pixels are in the same state. 
@@
 correspond to the maxima (resp. minima) of the blaze number. 
-Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.
+Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.\\
 
@@
   \label{fig:mu_vs_diff}
+\end{figure}
+
+\subsection{Experimental measure of the optimal incident angles}
+
+We experimentally characterized a DLP9500 DMD with a pixel pitch of $d = 10.8$\textmu m at $\lambda = 473$nm.
+The illumination geometry and DMD orientation are as presented in Fig.\ref{fig:2d_geom}.
+The experimental setup, shown in Fig.\ref{fig:setup_diff}, consists of an objective (L), a lens (L'), and a pinhole to produce a clean, collimated laser beam.
+A diaphragm is used to control the size of the beam incident on the DMD, and the reflected light is observed on a screen.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width = \textwidth]{paper/images/set_up_blaze.pdf}
+  \caption{
+    \textbf{Blazed number maxima measurement set up.
+    }
+    A laser beam is filtered and expanded using a telescope (objective L and lens L').
+A diaphragm is used to adjust the beam diameter before it reflects off the DMD with all pixels in \textit{on} (resp.
+\textit{off}) state.
+A movable screen allows to see the brightest diffraction orders as the DMD is rotated.
+  }
+  \label{fig:setup_diff}
+\end{figure}
+
+We then illuminate  with a collimated monochromatic beam the DMD in its full-\textit{on} (resp. full-\textit{off}) state, 
+corresponding to all pixels being rotated by an angle $\theta_B$ (resp. $-\theta_B$), 
+with $\theta_B=12$ degrees.
+The light distribution across the diffraction orders can be assessed directly with the naked eye by placing a screen after reflection from the DMD. 
+The DMD is mounted on a rotating plate with its axis of rotation orthogonal to the table, allowing variation of the angle $\alpha$, as defined in Fig.~\ref{fig:2d_geom}.
+The reference angle ($\alpha = 0$) is determined by auto-collimating the incident beam with the DMD in its idle state ($\theta_B = 0$).
+
+The DMD is then rotated until the reflected light is almost entirely on a single diffraction order, with adjacent orders extinguished. 
+The corresponding values of $\alpha$  are shown in Fig.\ref{fig:exp_vs_theo} as red lines, 
+illustrating a good correlation between the computed evolution of 
+the blazed number $\mu$ (Eq.\ref{eq:blazed_number}) 
+and the diffraction efficiency.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.8\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{images/theo_VS_exp_diffraction.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.8\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{images/theo_VS_exp_diffraction_off.pdf}
+  \end{subfigure}
+  \caption{
+    \textbf{Optimal angles for diffraction efficiency, experiment vs predictions.}
+     Blazing number $\mu$ as a function of the incident angle $\alpha$  (top)
+    for a pixel pitch of $d=7.6$\textmu m
+    and $\lambda = 473\text{nm}$, 
+    when all the pixels are \textit{on}-state (resp. \textit{off}-state), 
+    corresponding to a tilt angle $\theta=12^\circ$ (resp. $\theta=-12^\circ$).
+    Blue curves represent the predictions of Eq.(\ref{eq:blazed_number}) 
+    and red curves the experimentally measured angles for which the 
+    energy is maximal in one diffraction spot.
+  }
+  \label{fig:exp_vs_theo}
 \end{figure}
@@
 \end{figure}
-% For an incident angle $\alpha'$ 
+% For an incident angle $\alpha$ 
 % and a reflected angle $\beta'$ 
@@
 as a function of the wavelength for two pixel pitches $d=7.6$\textmu m and $d=10.8$\textmu m,
-with an incident angle of $\alpha' = 20^\circ$.
+with an incident angle of $\alpha = 20^\circ$.
 Within the visible spectrum, $\mu$ fluctuates between $0$ to $1$ over a typical range of roughly 100 nm.\\
@@
     Blazing number $\mu$ (Eq.(\ref{eq:blazed_number})) as a function of the incident wavelength
-    for an incident angle $\alpha' = 20^\circ$
+    for an incident angle $\alpha = 20^\circ$
     and a pixel pitch of $d=7.6$\textmu m (blue curve)
@@
 
-\section*{Appendix A: 1D calculation of the DMD diffraction effect}
-\label{app:1D}
-
-Assuming the effect of the device's finite size and illumination to be negligible,
-and situating ourselves within the context of the small-angle approximation,
-we can represent the field reflected from the device
-under the influence of plane wave illumination
-across two systems as follows:
-
-\begin{equation}
-  \begin{aligned}
-    R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
-    R_\text{blazed}(x) & \propto
-    \underbrace{
-      \left(
-      \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
-      \times
-    \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle}                                         \\ \text{+ angle of incidence}}}
-      \right)
-    }_\text{pixel response}
-    \otimes_x
-    \left[
-      \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
-      .
-      \underbrace{
-        e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
-      }_\text{angle of incidence}
-      \right]
-    \, ,
-  \end{aligned}
-  \label{eq:grating_response}
-\end{equation}
-
-with $\Pi(x)$ the rectangular function,
-representing the finite size of the pixel,
-defined as:
-
-\begin{equation}
-  \Pi(x) =
-  \begin{cases}
-    1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
-    0, & \text{otherwise}.
-  \end{cases}
-\end{equation}
-
-The intensity as a function of the angle in the far-field
-is given, up to a homotetic transformation,
-by the absolute value squared of
-
-The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
-
-\begin{equation}
-  \begin{aligned}
-    I_\text{flat}(\theta) \propto
-    \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
-    \times
-    \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
-    I_\text{blazed}(\theta) \propto
-    \underbrace{
-      \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D)
-    }_\text{orders of diffraction}
-    \times
-    \underbrace{
-      \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
-    }_\text{envelope} \, .
-  \end{aligned}
-\end{equation}
-
-We observe that the envelope (right hand term) is maximal for $\text{sin}(\theta_{max}) = \text{sin}(\alpha-2\theta_B)$
-while the effect of the periodicity (left hand term) is maximal for $\text{sin}(\theta_p)+ \text{sin}(\alpha) = p \lambda/d$,
-representing the orders of diffraction.
+% \section*{Appendix A: 1D calculation of the DMD diffraction effect}
+% \label{app:1D}
+
+% Assuming the effect of the device's finite size and illumination to be negligible,
+% and situating ourselves within the context of the small-angle approximation,
+% we can represent the field reflected from the device
+% under the influence of plane wave illumination
+% across two systems as follows:
+
+% \begin{equation}
+%   \begin{aligned}
+%     R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
+%     R_\text{blazed}(x) & \propto
+%     \underbrace{
+%       \left(
+%       \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
+%       \times
+%     \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle}                                         \\ \text{+ angle of incidence}}}
+%       \right)
+%     }_\text{pixel response}
+%     \otimes_x
+%     \left[
+%       \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
+%       .
+%       \underbrace{
+%         e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
+%       }_\text{angle of incidence}
+%       \right]
+%     \, ,
+%   \end{aligned}
+%   \label{eq:grating_response}
+% \end{equation}
+
+% with $\Pi(x)$ the rectangular function,
+% representing the finite size of the pixel,
+% defined as:
+
+% \begin{equation}
+%   \Pi(x) =
+%   \begin{cases}
+%     1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
+%     0, & \text{otherwise}.
+%   \end{cases}
+% \end{equation}
+
+% The intensity as a function of the angle in the far-field
+% is given, up to a homotetic transformation,
+% by the absolute value squared of
+
+% The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
+
+% \begin{equation}
+%   \begin{aligned}
+%     I_\text{flat}(\theta) \propto
+%     \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
+%     \times
+%     \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
+%     I_\text{blazed}(\theta) \propto
+%     \underbrace{
+%       \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D)
+%     }_\text{orders of diffraction}
+%     \times
+%     \underbrace{
+%       \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
+%     }_\text{envelope} \, .
+%   \end{aligned}
+% \end{equation}
+
+% We observe that the envelope (right hand term) is maximal for $\text{sin}(\theta_{max}) = \text{sin}(\alpha-2\theta_B)$
+% while the effect of the periodicity (left hand term) is maximal for $\text{sin}(\theta_p)+ \text{sin}(\alpha) = p \lambda/d$,
+% representing the orders of diffraction.
 
@@
 \noindent 
-S.M.P, M.W.M, and R.G.C and acknowledge the French \textit{Agence Nationale pour la Recherche} grants No. ANR-23-CE42-0010-01 MUFFIN and No. ANR-20-CE24-0016 MUPHTA 
+S.M.P, M.W.M, and R.G.C acknowledge the French \textit{Agence Nationale pour la Recherche} grants No. ANR-23-CE42-0010-01 MUFFIN and No. ANR-20-CE24-0016 MUPHTA 
 and the  Labex WIFI grant No. ANR-10-LABX-24, ANR-10-IDEX-0001-02 PSL*.
+
+L.M, J.C, M.G, and R.G acknowledge Cathie Ventalon for sharing a V-module and for her valuable discussions.
+L.M, J.C, M.G, and R.G were funded by Centre National de la Recherche Scientifique MITI/80Prime and the iXcore - iXlife - iXblue Fundation.
 

### 2025-12-05 paper/main.tex

@@
 \definecolor{light_orange}{RGB}{255,193,134}
-% \newcommand{\red}[1]{\textcolor{red}{#1}}
-\newcommand{\red}[1]{#1}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+% \newcommand{\red}[1]{#1}
 
@@
 
+% macro for red color
+
 %Uncomment next line if AMS fonts required
@@
 have emerged as powerful tools for wavefront shaping in complex media
-since the seminal work of A. Most and I. Vellekoop in the mid-2000s~\cite{Vellekoop2007focusing}.
+since the seminal work of A. \red{Mosk} and I. Vellekoop in the mid-2000s~\cite{Vellekoop2007focusing}.
 Nonetheless, LC-SLMs are hampered by their slow response time,
@@
 In particular, we provide characterization and validation procedures that require minimal changes compared to typical wavefront shaping setups. 
-\red{
 More specifically, we place ourselves in a standard experiment where the goal is to shape the complex wavefront impinging on a complex medium and control or measure its output response. 
 This is typically the case for a focusing experiment~\cite{Vellekoop2007focusing} or for measuring the transmission matrix of a complex medium~\cite{Popoff2010Measuring}. 
-}\\
+\red{
+In such works,
+complex modulation is usually achieved by encoding the optical phase
+into the spatial displacement of binary fringes
+displayed on the DMD,
+followed by filtering high spatial frequencies in the Fourier plane~\cite{lee1979binary}.
+Such a configuration permits multi-level complex modulation,
+at the cost of a reduced spatial resolution.
+The implementation and performance of such systems are discussed in detail in a separate tutorial~\cite{Gutierrez2024DMD}
+and are not elaborated further here.
+For the remainder of this paper, it will be assumed that the DMD is used
+for complex modulation via such a method.
+}
+\\
+
 We first introduce the diffraction properties of a DMD
@@
 
-\subsection{A 1D model}
-
 A significant distinction between liquid crystal modulators and DMDs
@@
 This difference gives rise to diffraction effects that can adversely affect
-both modulation quality and system efficiency.
+both modulation quality and diffraction efficiency, 
+\red{defined the fraction of the incident optical power that 
+is redistributed into a single diffraction order.}.
 The impact of these diffraction effects is highly dependent on several factors:
@@
 ranging approximately from $5$ to $\sim25$ \textmu m~\cite{TI}.\\
+
+\subsection{A 1D model}
 
@@
 This translates to an effective active pixel size of $d' < d$.\\
+\red{
+We compare here two types of gratings:
+Flat gratings, corresponding to LC-SLMs (Liquid Crystal Spatial Light Modulators),
+in which all pixels are in the same state;
+and blazed gratings, corresponding to DMDs,
+in which all pixels are tilted by the same angle.
+A grating gives rise to multiple diffraction orders,
+whose amplitudes are modulated by a global envelope.
+Importantly, these two effects can be decoupled.
+The diffraction order angles $\theta_p$
+are determined by the grating periodicity,
+while the envelope, and in particular the position of its maximum $\theta_{\mathrm{envelope}}^{\max}$,
+is governed by the structure of a single pixel,
+which constitutes the unit cell of the grating.
+}
+
+\subsubsection{Diffraction orders\\}
 
@@
 and angles $\theta_p$, as dictated by the grating equation:
-$
+\begin{equation}
     \text{sin}(\theta_p)+ \text{sin}(\alpha) = p\lambda/d = p \, \text{sin}(\theta_D),    
-$
+    \label{eq:1D_orders}
+\end{equation}
+
 where $\lambda$ is the wavelength of the light, $\alpha$ is the incident angle, 
-$\theta_D = \arcsin \left(\lambda/S\right)$ is the angle of the first diffraction order,
+$\theta_D = \arcsin \left(\lambda/d\right)$ is the angle of the first diffraction order,
 and $p$ is an integer value denoting the orders of diffraction.
@@
 and that is governed by $d$, $d'$, and its rotation angle in the case of a DMD.
-Importantly, we can decouple the effects of the grating's periodicity,
-which influences the angles of the diffraction orders,
-from the effects of the response of a single pixel,
-which shapes the envelope of the angular response.\\
+\red{
+While this regime is not always relevant,
+we can adopt the small-angle approximation to obtain
+a qualitative picture of the effect of the incident angle $\alpha$
+on the diffraction orders.
+In this case, we have $\theta_p \approx p\lambda/d - \alpha$.
+Thus, changing the incident angle effectively
+rotates all diffraction orders by an angle $-\alpha$.
+}\\
+% Importantly, we can decouple the effects of the grating's periodicity,
+% which influences the angles of the diffraction  \red{$\theta_p$},
+% from the effects of the response of a single pixel,
+% which shapes the envelope of the angular response.\\
 
@@
 
-For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
-we can assume that the response of a single pixel is uniform over its surface.
-Similarly, in the scenario of a blazed grating,
-such as a DMD, a linear phase slope is present on each pixel.
-This is due to the tilt angle $\theta_B$ of the mirrors.
-For an arbitrarily selected incidence angle $\alpha$,
-a global phase slope is introduced.
-This results in a trivial shift of the angular diffraction pattern by an angle $\alpha$.
-The incident angle $\alpha$ has the effect of shifting the entire angular diffraction pattern
-relative to the case of normal incidence,
-while the blazed angle $\theta_B$
-—the tilt angle of the micro-mirrors of the DMD—
-only shifts the envelope by an angle of $2\theta_B$.
-Whenever the fill factor approaches 100\%,
+\subsubsection{Position of the maximum of the envelope\\}
+% For a case of normal incidence with an LC-SLM (Liquid Crystal Spatial Light Modulator),
+% we can assume that the response of a single pixel is uniform over its surface.
+% Similarly, in the scenario of a blazed grating,
+% such as a DMD, a linear phase slope is present on each pixel.
+% This is due to the tilt angle $\theta_B$ of the mirrors.
+\red{
+The envelope of the diffracted field,
+which sets the amplitude of each diffraction order,
+is governed by the response of each unit cell, i.e. each pixel.
+In a flat grating excited at normal incidence,
+the envelope is peaked around the normal to the surface, i.e. $\theta_{envelope}^{max} = 0$.
+In a blazed grating,
+the tilt angle $\theta_B$ introduces a linear phase ramp
+in the response of each individual pixel.
+A flat grating can be seen as a particular case with $\theta_B = 0$.
+The incidence angle $\alpha$ also adds
+a global phase slope.
+When the fill factor approaches 100\%,
 i.e. when $d \approx d'$,
-the envelope for a flat grating achieves its maximum at $\theta_0 = -\alpha$;
-this corresponds to the angle of the zeroth-order diffraction,
-and the intensity is comparatively small for all other orders.
-In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
-The addition of a blazed angle $\theta_B$ results in both
-a shift in the envelope and in the position of its maximum 
-$\theta_0 = 2\theta_B - \alpha$.
-In the general case,
-this position may not align with a diffraction order anymore~\cite{Park2015properties}.
-We provide in Appendix A.  a 1D computation of this effect.
-A more accurate computation of the far field can be found  in ~\cite{Wang2023diffraction}.\\
-
-We represent in Fig.~\ref{fig:gratings} the angular response of a flat grating and a blazed grating
-for a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
+the cumulative effect leads to a shift of the envelope
+by an angle $2\theta_B - \alpha$.
+This results in a maximum of the envelope at
+}
+
+\begin{equation}
+    \theta_{envelope}^{max}= 2\theta_B - \alpha
+    \label{eq:1D_envelope}
+\end{equation}
+
+\red{
+which reduces to $\theta_{envelope}^{max} = -\alpha$ in the case of a flat grating.
+We detail the calculation of this effect in Appendix~A.
+}
+
+\subsubsection{Blazed grating condition\\}
+
+\red{
+In the case of a flat grating ($\theta_B = 0$), we see from Eqs.~\ref{eq:1D_orders} and~\ref{eq:1D_envelope}
+that the position of the first order is always aligned with the maximum of the envelope
+at the angle $-\alpha$,
+when the filling fraction is close to 1.
+In the general case $\theta_B \neq 0$,
+$\theta_{envelope}^{max}$ does not, in general, coincide
+with any diffraction order~\cite{Park2015properties, Wang2023diffraction}.
+This leads to a redistribution of the energy of the incoming beam
+into multiple diffraction orders,
+resulting in a reduced diffraction efficiency
+as well as cross-talk between different pixel states,
+as studied in section~\ref{sub:crosstalk}.\\
+}
+
+\red{
+The maximal diffraction efficiency is obtained 
+when the maximum of the envelope matches one order of diffraction. 
+This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}, 
+i.e. where there exist an integer $p$ so that we satisfy:
+}
+
+\begin{equation}
+  \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
+  = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
+  = p\frac{\lambda}{d} \, .
+  \label{eq:blazed_eq}
+\end{equation}
+
+\red{
+In particular, in the Littrow configuration, 
+i.e. for $\alpha = \theta_B$, we have $\theta_{max}^{envelope} = \alpha$; 
+the diffraction and incidence angles are identical. 
+The optimal condition is then satisfied for $\sin(\theta_B)={p\lambda}/{(2d)} = \sin(\theta_D)/2$.
+}
+
+% The later results in a trivial shift of the angular diffraction pattern by an angle $-\alpha$.
+% The incident angle $\alpha$ has the effect of shifting the entire angular diffraction pattern
+% relative to the case of normal incidence,
+% while the blazed angle $\theta_B$
+% —the tilt angle of the micro-mirrors of the DMD—
+% only shifts the envelope by an angle of $2\theta_B$.
+% Whenever the fill factor approaches 100\%,
+% i.e. when $d \approx d'$,
+% the envelope for a flat grating achieves its maximum at \red{$\theta_{envelope}^{max} = -\alpha$};
+% this corresponds to the angle of the zeroth-order  \red{$\theta_{p=0}$},
+% and the intensity is comparatively small for all other orders.
+% In these specific conditions, a singular diffraction order is visually perceived, corresponding to the optimum scenario.
+% The addition of a blazed angle $\theta_B$ results in both
+% a shift in the envelope and in the position of its maximum 
+% \red{$\theta_{envelope}^{max}= 2\theta_B - \alpha$. 
+% The envelope is peaked at the angle corresponding to the reflection on a 
+% unit cell, which consists in a mirror tilted with an angle $\theta_B$ 
+% with respect to the surface of the grating.}
+% \red{
+% We show in Appendix A. that the angles of the diffraction orders
+% are given by
+% $
+% \theta_p = \arcsin \left( 
+%     p\,\lambda/d -sin(\alpha) \right)
+% $,
+
+% }
+
+\subsubsection{Example\\}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width = 0.75 \textwidth]{paper/images/Fig3bis.pdf}
+  \caption{
+    \textbf{1D grating example.}
+    All angles are defined with respect to the normal of the surface of the the grating, 
+    denoted by $\vec{n}_G$, and are taken as positive for clockwise rotations and negative for counter-clockwise rotations.
+    The blazing angle $\theta_B$ is the angle between the normal to the micro-mirror $\vec{n}_M$ and $\vec{n}_G$.
+    The angles of the diffraction orders $\theta_p$ are independent of $\theta_B$, while the maximum of the envelope is at $\theta_{max}^{envelope} = 2\theta_B -\alpha$.
+  }
+  \label{fig:grating_geom_example}
+\end{figure}
+
+\red{
+We represent in Fig.~\ref{fig:grating_geom_example} an example of a geometry 
+for a 1D blazed grating and 
+in Fig.~\ref{fig:gratings} its angular response 
+as well as the one corresponding to a flat grating with the same parameters.}
+We take a 1D filling fraction of 95\% (correponding to a 2D filling fraction of $\approx 90$\%).
 For a flat grating, the zero-th order contains most of the intensity,
@@
 This state is achieved when the conditions of the blazed grating equation are fulfilled~\cite{Casini2014on}:
-
-\begin{equation}
-  \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
-  = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
-  = p\frac{\lambda}{d} \, .
-  \label{eq:blazed_eq}
-\end{equation}
+}
+
+% \begin{equation}
+%   \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
+%   = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
+%   = p\frac{\lambda}{d} \, .
+%   \label{eq:blazed_eq}
+% \end{equation}
+
+% \red{
+% In particular, in the Littrow configuration, i.e. for $\alpha = \theta_B$, 
+% we have $\theta_{max}^{envelope} = \alpha$; 
+% the diffraction and incidence angles are identical. 
+% The optimal condition is then satisfied for 
+% $\sin(\theta_B)={p\lambda}/{(2d)} = \sin(\theta_D)/2$.
+%  }
+% For the special case, where sin(theta_B)=p\lambda/d  that all the energy goes to the p order for \theta_b=\alpha.
+The intuition is that that for this wavelength a plane wave  accumulates phase of p*2*pi, so like a flat mirror.
 
@@
     \centering
-    \includegraphics[width = \textwidth]{images/gratings_flat.pdf}
+    \includegraphics[width = \textwidth]{images/flat_grating_lin_log.pdf}
     \label{fig:flat_grating}
@@
     \centering
-    \includegraphics[width = \textwidth]{images/gratings_blazed.pdf}
+    \includegraphics[width = \textwidth]{images/blazed_grating_lin_log.pdf}
     \label{fig:blazed_right}
@@
 
-To analyze more precisely the effect of diffraction in a DMD,
-one needs to consider the 2D surface of the modulator.
-We can establish a Cartesian coordinate system on the plane of the DMD,
-with axes $x$ and $y$ aligned with the pixel sides
-(refer to Fig.~\ref{fig:2d_geom}a).
-The pixels are uniformly repeated along these axes.
-However, a technical challenge arises
-in that the axis of rotation of the pixels
-aligns with the pixel diagonals,
-resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
-For the convenience of alignment and manipulation of the optical setup,
-it is preferable to work with the incident and outgoing beams
-which have the optical axis contained in the horizontal plane,
-i.e. a plane parallel to the table surface.
-A straightforward and common solution is to rotate the chip by 45 degrees
-relative to the horizontal plane,
-which aligns the pixel axis of rotation to be vertical.
-This configuration is depicted in Fig.~\ref{fig:2d_geom}b.\\
+% To analyze more precisely the effect of diffraction in a DMD,
+% one needs to consider the 2D surface of the modulator.
+% We can establish a Cartesian coordinate system on the plane of the DMD,
+% with axes $x$ and $y$ aligned with the pixel sides
+% (refer to Fig.~\ref{fig:2d_geom}a).
+% The pixels are uniformly repeated along these axes.
+% However, a technical challenge arises
+% in that the axis of rotation of the pixels
+% aligns with the pixel diagonals,
+% resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
+% For the convenience of alignment and manipulation of the optical setup,
+% it is preferable to work with the incident and outgoing beams
+% which have the optical axis contained in the horizontal plane,
+% i.e. a plane parallel to the table surface.
+% A straightforward and common solution is to rotate the chip by 45 degrees
+% relative to the horizontal plane,
+% which aligns the pixel axis of rotation to be vertical.
+\subsubsection{Blazed grating condition\\}
+\red{
+A DMD consists of a 2D array of micro-mirrors.
+In most DMDs, the rotation axis of the mirrors
+is aligned with the pixel diagonals.
+For convenience in the alignment and manipulation of the optical setup,
+it is preferable to work with incident and outgoing beams
+whose optical axes lie in the horizontal plane.
+A straightforward and common solution is to rotate the chip by $45^\circ$
+with respect to the horizontal plane,
+which makes the pixels' rotation axis vertical.
+This configuration is depicted in Fig.~\ref{fig:2d_geom}.
+We show in Appendix B. that it leads to a new blazed grating equation that reads:
+}\\
 
@@
 
-A more detailed description of the 2D system 
-and computation of its response 
-can be found in~\cite{Scholes2019structured}.
-However, we can find a simple condition for
-having most of the energy in one diffraction order 
-by looking at the effect of the tilts of the mirrors 
-when the axis is rotated by 45 degrees compared to 
-the axis of the pixel array (axes $x$ and $y$). 
-We consider a plane wave with an incoming angle $\alpha$ contained in the horizontal plane 
-(i.e. orthogonal to the rotation axis of the micro-mirrors).
-The effect of this angle is represented by a phase term of the form:
-
-\begin{equation}
-{\vec{k}_{in}\cdot\vec{r}} = {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}}
-\, ,
-\end{equation}
-
-with $\vec{k}_{in}$ the incoming wavevector and $\vec{r}$ the position in the modulator plane, 
-represented by the coordinate $(x,y)$.
-
-The grating condition is achieved when the output contributions are in phase 
-for the output angle corresponding to the specular reflection on the tilted mirrors. 
-It corresponds to the plot on the left in Fig.\ref{fig:gratings} for the 1D case.
-The phase contribution resulting from an output angle $2\theta_B-\alpha$, 
-corresponding to the specular reflection on each micro-mirror,
-reads:
-
-\begin{equation}
-\vec{k}_{out}\cdot\vec{r} 
-= {\frac{2\pi}{\lambda}\sin\left(2\theta_B - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}}
-\, ,
-\end{equation}
-
-The total phase then reads:
-
-\begin{equation}
-\phi(x,y) = \vec{k}_{in}.\vec{r}  + \vec{k}_{out}.\vec{r} =  
-{\frac{2\pi}{\lambda}\sin\left(2\theta - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}} +
-{\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}} 
-\, .
-\end{equation}
-
-We see that $x$ and $y$ have the same effect.
-The grating condition is then achieved when two consecutive pixels, 
-separated by a distance $d$ in both the $x$ and $y$ directions, are in phase.
-I.e. for
-
-\begin{equation}
-  \phi(x,y) = \phi(x+d,y) = \phi(x,y+d) \, .
-\end{equation}
-
-The new blazed grating equation then reads~:
+% A more detailed description of the 2D system 
+% and computation of its response 
+% can be found in~\cite{Scholes2019structured}.
+% However, we can find a simple condition for
+% having most of the energy in one diffraction order 
+% by looking at the effect of the tilts of the mirrors 
+% when the axis is rotated by 45 degrees compared to 
+% the axis of the pixel array (axes $x$ and $y$). 
+% We consider a plane wave with an incoming angle $\alpha$ contained in the horizontal plane 
+% (i.e. orthogonal to the rotation axis of the micro-mirrors).
+% The effect of this angle is represented by a phase term of the form:
+
+% \begin{equation}
+% {\vec{k}_{in}\cdot\vec{r}} = {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}}
+% \, ,
+% \end{equation}
+
+% with $\vec{k}_{in}$ the incoming wavevector and $\vec{r}$ the position in the modulator plane, 
+% represented by the coordinate $(x,y)$.
+
+% The grating condition is achieved when the output contributions are in phase 
+% for the output angle corresponding to the specular reflection on the tilted mirrors. 
+% It corresponds to the plot on the left in Fig.\ref{fig:gratings} for the 1D case.
+% The phase contribution resulting from an output angle $2\theta_B-\alpha$, 
+% corresponding to the specular reflection on each micro-mirror,
+% reads:
+
+% \begin{equation}
+% \vec{k}_{out}\cdot\vec{r} 
+% = {\frac{2\pi}{\lambda}\sin\left(2\theta_B - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}}
+% \, ,
+% \end{equation}
+
+% The total phase then reads:
+
+% \begin{equation}
+% \phi(x,y) = \vec{k}_{in}.\vec{r}  + \vec{k}_{out}.\vec{r} =  
+% {\frac{2\pi}{\lambda}\sin\left(2\theta - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}} +
+% {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}} 
+% \, .
+% \end{equation}
+
+% We see that $x$ and $y$ have the same effect.
+% The grating condition is then achieved when two consecutive pixels, 
+% separated by a distance $d$ in both the $x$ and $y$ directions, are in phase.
+% I.e. for
+
+% \begin{equation}
+%   \phi(x,y) = \phi(x+d,y) = \phi(x,y+d) \, .
+% \end{equation}
+
+% The new blazed grating equation then reads~:
 
@@
 
+\subsubsection{Blazed number\\}
 We can quantify how close we are to the ideal case,
-i.e., when satisfying the blazed this equation,
+i.e., when satisfying the blazed equation,
 by defining a {\em blazed number} $\mu$ as introduced in~\cite{WFSnet_diffraction}:
@@
 i.e. when one order of diffraction contains most of the energy,
-and minimal when we are in the worst-case scenario,
-i.e. when four orders of diffraction have a significant and equal intensity.\\
-
+\red{ 
+corresponding to one diffraction order $p$
+being close to the maximum of the envelope
+($\theta_p \approx \theta^{\mathrm{envelope}}_{\max}$).
+$\mu$ is minimal in the worst-case scenario,
+i.e. when four diffraction orders have a significant and equal intensity.
+Note that, since the maximum of the envelope is no longer aligned
+with the zeroth order, corresponding to the specular reflection on the micro-mirrors,
+the value of the order $p$ that aligns with the envelope peak
+when $\mu \approx 1$ can take any integer value,
+and depends on the DMD parameters and on the angle of incidence $\alpha$.
+}
+\\
+
+\subsubsection{Simulations\\}
 To demonstrate the effect, we first conduct a simulation of a DMD using Python
@@
 and the wavelength for pixel pitches of $d=7.6$\textmu m and $d=10.8$\textmu m.
-It should be noted that the efficiency of diffraction can be altered by adjusting
+It should be noted that the efficiency of diffraction,
+\red{i.e. the fraction of the incident optical power that is diffracted into a given diffraction order 
+relative to the incident power, }
+can be altered by adjusting
 the angle of incidence $\alpha$.
@@
 for an output angle $2\theta_B - \alpha$
-(marked by a yellow cross).
+(marked by a \red{red} cross).
 We see that for small positive values of $\alpha$,
@@
     for two incident angles $\alpha = -15^\circ$ and $\alpha = 0^\circ$.
-    The red cross indicates the maximum of the envelope $\theta_\text{max} = 2\theta_B - \alpha$.
+    The red cross indicates the maximum of the envelope $\theta^{envelope}_\text{max} = 2\theta_B - \alpha$
+    \red{with $\theta_B = -12^\circ$}.
   }
@@
     Color map representation of the blazed number for a pixel pitch of $d=7.6$\textmu m (a.)
-    and $d=10.8$\textmu m (b.).
+    and $d=10.8$\textmu m (b.) \red{for $\theta_B = -12^\circ$}.
   }
@@
 correspond to the maxima (resp. minima) of the blaze number. 
-Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.\\
+Note that the actual values of the minimum and maximum diffraction efficiency depend on the parameters of the DMD, such as pixel pitch, filling fraction, resolution, etc.
+\red{
+In a practical situation, pixels are not all in the same state:
+the configuration is a mix of \textit{on} and \textit{off} states
+corresponding to angles $\pm \theta_B$.
+The diffraction efficiency then also depends
+on the modulation scheme and on the specific parameters used for the modulation.
+This topic is addressed in detail in a separate tutorial
+focusing on complex modulation using binary amplitude modulators~\cite{Gutierrez2024DMD}.
+}\\
 
@@
     as a function of the wavelength. 
-    The diffraction efficiency is estimated by comparing the energy in a diffracted limited spot at the 
+    The diffraction efficiency is estimated by comparing the energy in a \red{diffraction} limited spot at the 
     location of the brightest diffraction order to the total diffracted energy.
@@
 \subsection{Modulation cross-talk}
+\label{sub:crosstalk}
 
@@
 Consequently, these effects are commonly overlooked and rarely documented in the information provided by the manufacturers. 
-Nonetheless, they are non-negligible, with deviations from a flat surface
-that typically spans several wavelengths~\cite{Brown2021multicolor}.\\
+\red{
+However, studies suggest that
+the residual stresses developed in the thin films
+during the manufacturing process of the DMDs
+induce non-negligible surface deformations~\cite{WEI2004375, Cao2001curvature, Brown2021multicolor}.
+}
+% Nonetheless, they are non-negligible, with deviations from a flat surface
+% that typically spans several wavelengths~\cite{Brown2021multicolor}.\\
 
@@
 and custom software.
-Furthermore, they also demand
+\red{
+Those approaches require
 a meticulously calibrated and stable setup.
-In this section, we introduce a straightforward method to characterize the aberrations using a lens and a camera.
+Moreover, once the phase correction map has been obtained,
+it must then be aligned with the pixel array
+in order to implement the phase correction on the DMD
+and compensate for aberrations.
+}
+In this section, we introduce a straightforward method for characterizing aberrations using a lens and a camera.
+\red{
+The goal is to \textit{in situ} compensate for the aberrations
+using a phase modulation scheme (Lee holograms),
+by iteratively optimizing the point spread function (PSF)
+at the focal plane of a lens. 
+The resulting phase correction map is the conjugate of the one
+induced by the surface deformation of the DMD,
+and can subsequently be used to correct aberrations
+with any desired complex modulation scheme.
+}
 This technique can be employed for any system that offers phase modulation,
@@
     (c) shows the ideal intensity pattern corresponding to the numerical aperture of the illumination setup,
-    i.e. its theoretical ideal point spread function.
-    (d) show the actual recorded intensity pattern corresponding to a distorted point spread function due to the DMD aberrations.
+    i.e. its theoretical ideal PSF.
+    (d) show the actual recorded intensity pattern corresponding to a distorted PSF due to the DMD aberrations.
   }
@@
 
-We assume the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}.
+We assume that the DMD is configured to deliver phase modulation~\cite{lee1979binary,Gutierrez2024DMD}. 
+\red{
+It is classically done using Lee holograms or similar approaches,
+where the optical phase is encoded in the local spatial displacement
+of a regular array of amplitude fringes displayed on the DMD.
+By filtering a single diffraction order,
+the resulting field retains only the phase modulation.
+Specific implementations and variants of such approaches are discussed in~\cite{Gutierrez2024DMD}.
+}
 This implies that the modulator can be divided into $N$ sections,
@@
     using a lens (L$_\text{2}$).
-    The intensity pattern, up to a homotetic transformation,
+    The intensity pattern, up to a \red{homothetic} transformation,
     represent the input excitation on the medium
@@
     and the sum of the intensity elsewhere.    
-    We both want to maximize the intensity of the main lobe of the point spread function
+    We both want to maximize the intensity of the main lobe of the PSF
     and minimize its side lobes.
@@
 to be maximal for the ideal correction of optical aberrations.
+\red{
+The approach is inspired by methods used for focusing through
+complex disordered media}~\cite{Vellekoop2007focusing}\red{,
+by effectively treating the aberrations of the DMD
+as equivalent to propagation through an unknown linear medium.
+For each degree of freedom
+(i.e. the phase on one group of pixels in Ref.~\cite{Vellekoop2007focusing},
+or the coefficient of one Zernike polynomial in this section),
+we scan different values and retain the one that
+maximizes the intensity at the target location.
+}
 We first  generate a mask, represented by a disk,
@@
 The exact size of this radius for a successful optimization is not critical
-and can be determined by approximating the dimensions of the ideal point spread function,
+and can be determined by approximating the dimensions of the ideal PSF,
 expressed as $r_0 \approx M \frac{\lambda}{2 NA}$,
@@
 This calculated sum is then divided by an analogous product, but with the complementary mask substituted in place of the original
-in order to minimize the side lobes of the point spread function.
+in order to minimize the side lobes of the PSF.
 We set initial parameter values as $a'_n = 0\,\, \forall n \in [0..N]$.
@@
 To estimate the quality of the correction,
-we compute the Strehl ratio of the point spread function.
+we compute the Strehl ratio of the PSF.
 It is defined as the maximum of the measured PSF divided
@@
     \textbf{Effect of Zernike polynomials on the optimization of the PSF.}
-    (a) Strehl ratio of the point spread function as a function of the number of Zernike polynomials used
+    (a) Strehl ratio of the PSF as a function of the number of Zernike polynomials used
     for the compensation of the aberrations.
-    (b) Intensity profile of the point spread function for different number of Zernike polynomials
+    (b) Intensity profile of the PSF for different number of Zernike polynomials
     in linear and logarithmic scale.
@@
   This can be counteracted in situ using a standard setup and a straightforward optimization procedure
-  to maximize the intensity at the central position of the point spread function.
+  to maximize the intensity at the central position of the PSF.
 \end{tldr}
@@
 
-A simple yet effective solution consists in dampening the vibrations
+A simple yet effective solution consists in damping the vibrations
 at the flat cable's level by clamping it with a soft material,
@@
   \caption{
-    \textbf{Dumping mechanical vibrations.}
+    \textbf{\red{Damping} mechanical vibrations.}
     Pictures of experimental setups where a damping of the vibrations is implemented.
@@
 and the temporal averaging over the measured frames.
-Fig.~\ref{fig:temp_decorr} shows the measured decorrelation
+Figure~\ref{fig:temp_decorr} shows the measured decorrelation
 over time when running an sequence in an infinite loop.
@@
 to the start of a sequence
-and is set in the second one to
-4.2 hours post the after of the sequence (b).
-For the scenario (a), it is to note that before the sequence commenced,
+% and is set in the second one to
+% 4.2 hours post the after of the sequence (b).
+\red{and in the second case (b) is set to
+4.2 hours after the start of the sequence.}
+For the scenario (a), it is worth noting that before the sequence commenced,
 the DMD was active but remained in the idle state,
@@
 to ensure a good modulation quality and diffraction efficiency.
-Furhtermore, the DMD surface is not flat and can introduce aberrations
+\red{Furthermore}, the DMD surface is not flat and can introduce aberrations
 which can be corrected using a simple optimization procedure.
@@
 
-% \section*{Appendix A: 1D calculation of the DMD diffraction effect}
-% \label{app:1D}
-
-% Assuming the effect of the device's finite size and illumination to be negligible,
+\section*{Appendix A: 1D calculation of the DMD diffraction effect}
+\label{app:1D}
+
+Assuming the effect of the device's finite size and illumination to be negligible,
 % and situating ourselves within the context of the small-angle approximation,
-% we can represent the field reflected from the device
-% under the influence of plane wave illumination
-% across two systems as follows:
-
-% \begin{equation}
-%   \begin{aligned}
-%     R_\text{flat}(x)   & \propto \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) x} \\
-%     R_\text{blazed}(x) & \propto
-%     \underbrace{
-%       \left(
-%       \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
-%       \times
-%     \underbrace{e^{j\frac{2\pi}{\lambda}2\text{sin}(\theta_B-\alpha) x}}_{\substack{\text{blazed angle}                                         \\ \text{+ angle of incidence}}}
-%       \right)
-%     }_\text{pixel response}
-%     \otimes_x
-%     \left[
-%       \underbrace{\sum_k \delta(x-k d) }_\text{periodicitiy}
-%       .
-%       \underbrace{
-%         e^{-j\frac{2\pi}{\lambda}\text{sin}(\alpha) kd}
-%       }_\text{angle of incidence}
-%       \right]
-%     \, ,
-%   \end{aligned}
-%   \label{eq:grating_response}
-% \end{equation}
-
-% with $\Pi(x)$ the rectangular function,
-% representing the finite size of the pixel,
-% defined as:
-
-% \begin{equation}
-%   \Pi(x) =
-%   \begin{cases}
-%     1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
-%     0, & \text{otherwise}.
-%   \end{cases}
-% \end{equation}
+we can represent the field reflected from the device
+under the influence of plane wave illumination
+across two systems as follows.
+Here, $\theta_B$ denotes the blaze angle, i.e. the tilt of the micromirror normal
+with respect to the device normal, so that the corresponding specular direction is
+$2\theta_B-\alpha$:
+
+\begin{equation}
+  \begin{aligned}
+    R_\text{flat}(x)   & \propto
+    \left[\Pi\left(x/d'\right) \otimes_x \sum_k \delta(x-k d)\right]
+    e^{j\frac{2\pi}{\lambda}\sin(\alpha) x} \,,\\
+    R_\text{blazed}(x) & \propto
+    \underbrace{
+      \left(
+        \underbrace{\Pi\left(x/d'\right)}_\text{pixel size}
+        \,
+        \underbrace{e^{j \frac{2\pi}{\lambda}\left( \sin(2\theta_B-\alpha) - \sin(\alpha) \right) x}}_{\substack{\text{blazed angle}  \, , \\ \text{+ angle of incidence}}}
+      \right)
+    }_\text{pixel response}
+    \otimes_x
+    \underbrace{\sum_k \delta(x-k d) }_\text{periodicity}
+    \,
+    e^{j\frac{2\pi}{\lambda}\sin(\alpha) x}
+    \, .
+  \end{aligned}
+  \label{eq:grating_response}
+\end{equation}
+
+with $\Pi(x)$ the rectangular function,
+representing the finite size of the pixel,
+defined as:
+
+\begin{equation}
+  \Pi(x) =
+  \begin{cases}
+    1, & \text{if } -\frac{1}{2} < x < \frac{1}{2}, \\
+    0, & \text{otherwise}.
+  \end{cases}
+\end{equation}
+
+The filling fraction, in the 1D case, is the ratio of the size of the pixels (or micro-mirrors) $d'$,
+and the pitch $d$, such that $d'=\rho d$.\\
 
 % The intensity as a function of the angle in the far-field
-% is given, up to a homotetic transformation,
+% is given, up to a homothetic transformation,
 % by the absolute value squared of
-
-% The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
-
-% \begin{equation}
-%   \begin{aligned}
-%     I_\text{flat}(\theta) \propto
-%     \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha))
-%     \times
-%     \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right) \\
-%     I_\text{blazed}(\theta) \propto
-%     \underbrace{
-%       \sum_p \delta(\text{sin}(\theta)+\text{sin}(\alpha)-p\,\theta_D)
-%     }_\text{orders of diffraction}
-%     \times
-%     \underbrace{
-%       \text{sinc}^2\left( \left[\text{sin}(\theta)+\text{sin}(\alpha)\right] \frac{\lambda}{d'}\right)
-%     }_\text{envelope} \, .
-%   \end{aligned}
-% \end{equation}
+The Fourier transform of Eq.~\ref{eq:grating_response} can be written as:
+
+\begin{equation}
+  \begin{aligned}
+    I_\text{flat}(\theta) \propto
+      \sum_p \delta\bigl(\sin(\theta)+\sin(\alpha)-p\,\sin(\theta_D)\bigr)
+    \times
+    \operatorname{sinc}^2\left( \pi \rho\frac{
+        \sin(\theta)-\sin(\alpha)
+        }{
+        \sin(\theta_D)
+        }\right) \\
+    I_\text{blazed}(\theta) \propto
+    \underbrace{
+      \sum_p \delta\bigl(\sin(\theta)+\sin(\alpha)-p\,\sin(\theta_D)\bigr)
+    }_\text{orders of diffraction}
+    \times
+    \underbrace{
+      \operatorname{sinc}^2\left( \pi \rho\frac{
+        \sin(\theta)-\sin(2\theta_B - \alpha)
+        }{
+        \sin(\theta_D)
+        }\right)
+    }_\text{envelope} \, .
+  \end{aligned}
+\end{equation}
+
+We observe that the envelope (right-hand term) is maximal for
+$\sin(\theta_\text{max}) = \sin(2\theta_B - \alpha)$,
+while the effect of the periodicity (left-hand term) is maximal when
+$\sin(\theta_p) + \sin(\alpha) = p\,\lambda/d$,
+representing the orders of diffraction.
 
@@
 
+\section*{Appendix B: 2D calculation of blazed grating condition}
+
+To analyze more precisely the effect of diffraction in a DMD,
+one needs to consider the 2D surface of the modulator.
+We can establish a Cartesian coordinate system on the plane of the DMD,
+with axes $x$ and $y$ aligned with the pixel sides
+(refer to Fig.~\ref{fig:2d_geom}a).
+The pixels are uniformly repeated along these axes.
+However, a technical challenge arises
+in that the axis of rotation of the pixels
+aligns with the pixel diagonals,
+resulting in a rotation by 45 degrees with respect to the $x$ and $y$ axes.
+For the convenience of alignment and manipulation of the optical setup,
+it is preferable to work with the incident and outgoing beams
+which have the optical axis contained in the horizontal plane,
+i.e. a plane parallel to the table surface.
+A straightforward and common solution is to rotate the chip by 45 degrees
+relative to the horizontal plane,
+which aligns the pixel axis of rotation to be vertical.
+
+A more detailed description of the 2D system 
+and computation of its response 
+can be found in~\cite{Scholes2019structured}.
+However, we can find a simple condition for
+having most of the energy in one diffraction order 
+by looking at the effect of the tilts of the mirrors 
+when the axis is rotated by 45 degrees compared to 
+the axis of the pixel array (axes $x$ and $y$). 
+We consider a plane wave with an incoming angle $\alpha$ contained in the horizontal plane 
+(i.e. orthogonal to the rotation axis of the micro-mirrors).
+The effect of this angle is represented by a phase term of the form:
+
+\begin{equation}
+{\vec{k}_{in}\cdot\vec{r}} = {\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}}
+\, ,
+\end{equation}
+
+with $\vec{k}_{in}$ the incoming wavevector and $\vec{r}$ the position in the modulator plane, 
+represented by the coordinate $(x,y)$.
+
+The grating condition is achieved when the output contributions are in phase 
+for the output angle corresponding to the specular reflection on the tilted mirrors. 
+It corresponds to the plot on the left in Fig.\ref{fig:gratings} for the 1D case.
+The phase contribution resulting from an output angle $2\theta_B-\alpha$, 
+corresponding to the specular reflection on each micro-mirror,
+reads:
+
+\begin{equation}
+\vec{k}_{out}\cdot\vec{r} 
+= {\frac{2\pi}{\lambda}\sin\left(2\theta_B - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}}
+\, ,
+\end{equation}
+
+The total phase then reads:
+
+\begin{equation}
+\phi(x,y) = \vec{k}_{in}.\vec{r}  + \vec{k}_{out}.\vec{r} =  
+{\frac{2\pi}{\lambda}\sin\left(2\theta - \alpha\right) \frac{\left(x - y\right)}{\sqrt{2}}} +
+{\frac{2\pi}{\lambda}\sin\alpha \frac{\left(x - y\right)}{\sqrt{2}}} 
+\, .
+\end{equation}
+
+We see that $x$ and $y$ have the same effect.
+The grating condition is then achieved when two consecutive pixels, 
+separated by a distance $d$ in both the $x$ and $y$ directions, are in phase.
+I.e. for
+
+\begin{equation}
+  \phi(x,y) = \phi(x+d,y) = \phi(x,y+d) \, .
+\end{equation}
+
+The new blazed grating equation then reads~:
+
+\begin{equation}
+  \text{sin}(2\theta_B-\alpha) + \text{sin}(\alpha)
+  = 2 \,\text{sin}(\theta_B)  \,\text{cos}(\theta_B-\alpha)
+  = p\frac{\sqrt{2}\lambda}{d} \, ,
+  \label{eq:blazed_eq_new}
+\end{equation}
+
 \section*{Acknowledgments}
